### 2012-03-15 Examples/OPT/StatPDE/Example3/content.tex
Initial Transfer From OLD CVS Reprository

In this example, we solve a optrimization problems from 
fluid dynamcis. The configuration is similar to 
the fluid optimization problem proposed by Roland Becker 
in 2000 ``Mesh adaption for stationary flow control''.

We describe the underlying equations later...