### 2012-03-15 Examples/PDE/StatPDE/Example6/content.tex

todo. short: this example shows how to use the constraintsmaker class by enforcing periodic boundary conditions in 2d.