### 2012-03-16 Examples/Documentation/examples.tex
More structure for the documentation. And some more information on software requirements

\chapter{Example Handling, Creating new Examples}
\label{chap:howtoex}
To implement new examples or to use existing examples 
from the library for own research, the user 
can simply copy an existing example. In this 
new example, own code and changes can be compiled.

Adding a new regression test to the repository 
needs some concentration of the user. In the 
following, we explain how to do this and what 
\textit{must} be taken into account:
\begin{verbatim}
cp -r Example1 ExampleNew
cd ExampleNew
rm -r CVS
cd Test
rm -r CVS
\end{verbatim}
It is important to remove the repository information,
which is stored in the directory CVS.

Then, we have to consider the following step. In order to 
run the regression tests, we start the executable file
of the present example. This executable comes from the 
Makefile. For example in the Makefile of 
Example \ref{PDE_Stat_Stokes}, 
\begin{verbatim}
target   = $(BINDIR)/DOpE-PDE-StatPDE-Example1-$
           (dope_dimension)d-$(deal_II_dimension)d
\end{verbatim}
and concretely:
\begin{verbatim}
DOpE-PDE-StatPDE-Example1-2d-2d
\end{verbatim}
The aforementioned expression is the name of 
the executable that is required to run the 
regression test. Hence, in 
\begin{verbatim}
PDE/StatPDE/Example1/Test> emacs test.sh
\end{verbatim}
you find two times the line
\begin{verbatim}
echo "Running Program 
     ../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm"
     (../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm 2>&1) 
     > /dev/null
\end{verbatim} 
It is really important when copying an existing example,
to change this information. 
First, we have to change the target in the Makefile
\begin{verbatim}
target   = $(BINDIR)/DOpE-PDE-StatPDE-ExampleNew-$
           (dope_dimension)d-$(deal_II_dimension)d
\end{verbatim}
Second, we have to replace 
Example1 by ExampleNew such that 
\begin{verbatim}
echo "Running Program 
     ../../../../../bin/DOpE-PDE-StatPDE-ExampleNew-2d-2d test.prm"
     (../../../../../bin/DOpE-PDE-StatPDE-ExampleNew-2d-2d test.prm 2>&1) 
     > /dev/null
\end{verbatim} 
Please, also change the dimension parameters 
\begin{verbatim}
-2d-2d
\end{verbatim} 
Now, the user is prepared to change any information in the 
\begin{verbatim}
main.cc, localpde.h, functionals.h, localfunctional.h, etc
\end{verbatim}