### 2012-03-16 Examples/description.tex

\documentclass[a4paper,cleardoubleempty]{scrreprt}
\usepackage{amsmath,amssymb}
\usepackage[pdftex,bookmarks=false,pdfpagelabels=true,plainpages=false]{hyperref}
\usepackage{booktabs}
\usepackage{makeidx}
\usepackage{url}
%\usepackage{sfmath}
\newtheorem{Problem}{Problem}[section]

\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{black}{TODO: #1}}}}

\makeindex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title{DOpE - Deal Optimization Environment}
\author{
  Michael Geiger\thanks{University of Heidelberg, 
    {michael.geiger@iwr.uni-heidelberg.de}}, 
  Christian Goll\thanks{University of Heidelberg, 
    {christian.goll@iwr.uni-heidelberg.de}}, 
  Thomas Wick\thanks{University of Heidelberg, 
    {thomas.wick@iwr.uni-heidelberg.de}}, 
  Winnifried Wollner\thanks{University of Hamburg, 
    {winnifried.wollner@math.uni-hamburg.de}}
}

\date{\today}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\maketitle
\cleardoublepage\pagestyle{headings}
\tableofcontents
\cleardoublepage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Documentation/foreword.tex}
\input{Documentation/introduction.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Documentation/examples.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Examples for PDE Solution}
\label{PDE}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Stationary PDEs} 
\label{PDE_Stat}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stationary Stokes Equations} 
\label{PDE_Stat_Stokes}
\input{PDE/StatPDE/Example1/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stationary FSI with INH Material} 
\label{PDE_Stat_FSI_INH}
\input{PDE/StatPDE/Example2/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stationary FSI with STVK Material} 
\label{PDE_Stat_FSI_STVK}
\input{PDE/StatPDE/Example3/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stationary Elasticity Benchmark} 
\label{PDE_Stat_Elasticity}
\input{PDE/StatPDE/Example4/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stationary Plasticity Benchmark} 
\label{PDE_Stat_Plasticity}
\input{PDE/StatPDE/Example5/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stationary Stokes Equations with periodic BC} 
\label{PDE_Stat_Periodic_Stokes}
\input{PDE/StatPDE/Example6/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Laplace Equation in 2D} 
\label{PDE_Stat_Laplace_2D}
\input{PDE/StatPDE/Example7/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Laplace Equation in 3D}
\label{PDE_Stat_Laplace_3D}
\input{PDE/StatPDE/Example8/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Nonstationary PDEs}
\label{PDE_Instat}
\input{PDE/InstatPDE/content.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nonstationary Stokes Equations}
\label{PDE_Instat_Stokes}
\input{PDE/InstatPDE/Example1/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nonstationary FSI Problem}
\label{PDE_Instat_FSI}
\input{PDE/InstatPDE/Example2/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Black-Scholes Equation}
\label{PDE_Instat_Black_Scholes}
\input{PDE/InstatPDE/Example3/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Heat Equation in 1D}
\label{PDE_Instat_Heat_1D}
\input{PDE/InstatPDE/Example4/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Heat Equation in 2D with nonlinearity}
\label{PDE_Instat_Heat_2D}
\input{PDE/InstatPDE/Example5/content.tex}
\cleardoublepage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Examples with Optimization}
\label{OPT}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Subject to a Stationary PDE}
\label{OPT_Stat}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Distributed control with a linear elliptic PDE}
\label{OPT_Stat_Distrib_Lin_Ellipt}
\input{OPT/StatPDE/Example1/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parameter control with a linear elliptic PDE}
\label{OPT_Stat_Param_Lin_Ellipt}
\input{OPT/StatPDE/Example2/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parameter control with a nonlinear PDE from fluid dynamcis}
\label{OPT_Stat_Param_Nonlin_Fluid}
\input{OPT/StatPDE/Example3/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Control in the dirichlet boundary values}
\label{OPT_Stat_Dirichlet_Boundary}
\input{OPT/StatPDE/Example4/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Distributed Control with Different Meshes for Control and State}
\label{OPT_Stat_Distributed_MultiMesh}
\input{OPT/StatPDE/Example5/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Compliance Minimization of a variable Thickness MBB-Beam}
\label{OPT_Stat_MBB-Beam}
\input{OPT/StatPDE/Example6/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Distributed control with a linear elliptic PDE using SNOPT}
\label{OPT_Stat_Box_controlconstraints_SNOPT}
\input{OPT/StatPDE/Example7/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Topology optimization of an MBB-Beam using SNOPT}
\label{OPT_Stat_TopOpt_MBB_SNOPT}
\input{OPT/StatPDE/Example8/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parameter control with a non-linear PDE from FSI dynamics}
\label{OPT_Stat_Param_Nonlin_FSI}
%\input{OPT/StatPDE/Example5/content.tex}
\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Documentation/tests.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addcontentsline{toc}{chapter}{Index}
\printindex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}