### 2012-04-16 Examples/PDE/StatPDE/Example8/content.tex

\subsubsection{General problem description}
Laplace in 3D
on locally refined grids with Kelly estimator
with iterative solver (CG without preconditioning)