### 2012-04-17 Examples/PDE/StatPDE/Example9/content.tex
Added some documentation regarding the DWR-Example.

\subsubsection{General problem description}
This example shows the use of the adaptive grid refinement and error estimation by the \textit{DWR method} (For a description of the method, see \cite{BR03}.) applied to the laplace equation 
\begin{align*}
-\Delta u &= f && \text{ in } \Omega
\end{align*}
with the analytical solution
\begin{align*}
u = \sin\left(\frac{\pi}{x^2 + y^2}\right),
\end{align*}
 the corresponding right hand side $f = -\Delta u$ and appropriate Dirichlet Conditions on $\partial \Omega$, where the domain is given by
\begin{align*}
\Omega = [-2,2]^2\setminus \overline{B}_{0.5}(0).
\end{align*}
We want to estimate the error in the following functional of interest
$$\fullfunction{J}{H^1(\Omega)}{\mathbb{R}}{u}{\frac 1 {\abs{\Gamma}} \int_\Gamma u \,dx}$$
where
$\Gamma = \left\{(x,y)\in \mathbb R ^2\,\big | \,x=0,-2<y<0.5 \right \}$.

For this setting, we have the error representation
\begin{align}\label{dwr.error_identity}
J(e)=\sum_{K\in \mathbb T _h} \left\{(R_h, z-\psi_h)_K + (r_h, z-\psi_h)_{\partial_K}\right\}
\end{align}
with the error $e=u-u_h$, the Triangulation $\mathbb T_h$, the dual solution $z$, arbitraty function $\phi_h \in V_h$ (the ansatz space) and the cell- and edge-residuals: 
\begin{align}\label{dwr.cell_residual}
\eval[1]{R_h}_{K} = f+\Delta u_h
\end{align}
resp. 
\begin{align}\label{dwr.edge_residual}
\eval[1]{r_h}_{\Sigma} =
\begin{cases}
\frac 12 [\partial_n u_h] ,&\text{if } \Sigma \subset \partial_K \setminus \partial \Omega,\\
0,& \text{if } \Sigma \subset\partial \Omega.
\end{cases}
\end{align}

It holds $J(u) \approx 0.441956231972232$.
\subsubsection{Program description}
In this section we want to focus on what you have to do if you want to enhance your existing code to use the \textit{DWR method}.

First, additionally to all the things one has to do when just solving the equation, we have to include the file
\begin{verbatim}
higher_order_dwrc.h
\end{verbatim}

As we approximate the so called 'weights' $z-\phi_h$ in the error representation by a patchwise higher order interpolation of $z_h$ (the computed dual solution), we have to enforce patch-wise refinement of the grid by giving the flag 
\begin{verbatim}
Triangulation<2>::MeshSmoothing::patch_level_1
\end{verbatim}
to the triangulation.

To be able to solve the adjoint equation for the error estimation one needs to implement some methods regarding the equation as well es the functional of interest:
\begin{itemize}
\item In {\verb pdeinterface.h }
\begin{itemize}
\item {\verb CellEquation_U }: Weak form of the adjoint equation.
\item {\verb CellMatrix_T }: The FE matrix for the adjoint problem.
\item {\verb FaceEquation_U }: This one is needed in this case here because we have a functional of interest that lives on faces. 
\end{itemize}
\item {\verb functionalinterface.h }
\begin{itemize}
\item {\verb FaceValue_U }: This is the right hand side of the adjoint equation.
\end{itemize}
\end{itemize}
During the  evaluation of \eqref{dwr.error_identity}, the following methods are needed
\begin{itemize}
\item {\verb StrongCellResidual }: The cell residual, see \eqref{dwr.cell_residual}.
\item {\verb StrongFaceResidual }: The terms in \eqref{dwr.error_identity} that lies in the interiour (i.e. the jumps).
\item {\verb StrongBoundaryResidual }:The terms in \eqref{dwr.error_identity} that lies on the boundary (There are none in this case).
\end{itemize}

After this, we tell the problem which functional we want to use for the error estimation, this is done via
\begin{verbatim}
P.SetFunctionalForErrorEstimation(LFF.GetName())
\end{verbatim}
where \texttt{P} is of type \texttt{PDEProblemContainer} and \texttt{LFF} is the desired functional of interest.

The next thing we need is an object of the type
\begin{verbatim}
HigherOrderDWRContainer
\end{verbatim}
This container takes care of the computation of the weights.

To build this, we need the following:
\begin{itemize}
\item {\verb DOFH_higher_order }: With some higher order Finite Elemnts and the already defined triangulation, we build this \texttt{SpaceTimeHandler}. This is needed because  we want to use the patch-wise higher order interpolation of the weights.
\item {\verb idc_high }: A \texttt{IntegratorDataContainer} in which we put some (face)quadrature formulas for the evaluation of the error Identity.
\item A string which indicates how we want to store the weight-vectors (here: { \verb "fullmem" }).
\item \texttt{pr}: The \texttt{ParameterReader} which we have already defined.
\item A enum of type \texttt{EETerms} that tells the container, which error terms we want to compute (primal error indicators vs. dual error indicators, see \cite{BR03}).  
\end{itemize}
The last preparation step is now to initialize the \texttt{DWRDataContainer} with the problem in use:
\begin{verbatim}
  solver.InitializeHigherOrderDWRC(dwrc);
\end{verbatim}

Succeding the solution of the state equation 
\begin{verbatim}
solver.ComputeReducedFunctionals(); 
\end{verbatim}
we compute the error indicators by calling 
\begin{verbatim}
solver.ComputeRefinementIndicators(dwrc);
\end{verbatim}
We can now get the error indicators (with signs!) out of \texttt{dwrc} by 
\begin{verbatim}
dwrc.GetErrorIndicators();
\end{verbatim}
With these indicators, we are now able to refine our grid adaptively (there are several mesh adaption strategies implemented, like 'optimized', 'fixednumber' or 'fixedfraction')
\begin{verbatim}
DOFH.RefineSpace("optimized", &error_ind);
\end{verbatim}
Note, that one has to take the norm of each entry in the vector of the error indicators before feeding them into the \texttt{RefineSpace} method.