### 2012-05-05 Examples/OPT/InstatPDE/Example1/content.tex
Added an initial Example for the OPT-Instat case, it is not yet working, but will be soon...

\subsubsection{General problem description}

This example is a modified version of \texttt{PDE/InstatPDE/Example5}. 
Again, we consider the heat equation, this time with an additional nonlinear term
\begin{equation*}
\partial_t u(t,x,y) - \Delta u(t,x,y) + u(t,x,y)^2 = f(t,x,y),
\end{equation*}
with homogeneous dirichlet-data.
The computational domain is $\Omega \times I = [0,\pi]^2 \times [0,1]$. From the known solution, we can compute the appropriate data 
\begin{align*}
f(t,x,y) &= (3-2t)e^{t-t^2} \sin(x) \sin(y) + e^{(t-t^2)^2} \sin^2(x) \sin^2(y),\\
u_0(x,y) &= q(x,y).
\end{align*}

With the cost functional 
\[
 \min_{q,u} J(q,u) = \frac{1}{2}\int_{\Omega} (u(1,x,y) - \sin(x) \sin(y))^2\,d(x,y) + \frac{1}{2} \int_{\Omega} (q(x,y) - \sin(x) \sin(y))^2\,d(x,y).
\]
It has the optimal solution 
\begin{align*}
\overline{u}(t,x,y) &= e^{t-t^2} \sin(x) \sin(y),\\
\overline{q}(t,x,y) &= \sin(x) \sin(y)
\end{align*}
and $J(\overline{q},\overline{u}) = 0$.