### 2012-05-07 Examples/OPT/StatPDE/Example2/content.tex
OPT StatPDE Example2 uses now a pointfunctional as cost functional.

\subsubsection{General problem description}
This example solves the following pointwise minimization problem
\begin{align*}
\min_{(q,u)\in \mathbb R^3 \times H_0^1(\Omega; \mathbb R^2)} J(q,u) &= \frac{1}{2} \sum_{i=0}^2\abs{(u-\overline u)(x_i)}^2 + \frac{\alpha}{2}\|q\|^2\\
\text{s.t.} (\nabla u,\nabla \phi) &= (f(q),\phi)\;\;\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)
\end{align*}
on the domain $\Omega = [0,1]^2$, with
\begin{itemize}
\item the observation points
\begin{align*}
x_0 = (0.5, 0.5), \quad x_1 = (0.5, 0.25),\quad x_2 = (0.25, 0.25),
\end{align*}
\item the regularization parameter $\alpha = 0$, 
\item the right hand side
\begin{align*}
 f(q) &= q_0 \left(\begin{matrix}2\pi^2  \sin( \pi x) \sin(\pi y)\\0 \end{matrix}\right)\\
      &+ q_1 \left(\begin{matrix}5\pi^2  \sin( \pi x) \sin(2\pi y)\\0 \end{matrix}\right)\\
      &+ q_2 \left(\begin{matrix}0 \\8\pi^2  \sin(2\pi x) \sin(2\pi y)\end{matrix}\right)\\
\end{align*}
\item and the exact solution given by 
\begin{align*}
 \overline{q} &= \bigl(1;0.5;1\bigr)\\
 \overline{u}& = \left(\begin{matrix} \sin( \pi x)( \sin(\pi y)+0.5\sin(2\pi y))\\\sin(2\pi x) \sin(2\pi y) \end{matrix}\right).
\end{align*}
\end{itemize}
\subsubsection{Program description}
Notice that the cost functional is of mixed type (from our computational point of view), i.e. the first part is a pointfunctional whereas the regularization part requires the evaluation of a domain integral. So our \texttt{LocalFunctional} returns as his type:
\begin{verbatim}
      string
      GetType() const
      {
        return "point domain";
      }
\end{verbatim}
This brings along that we have not only to implement 
\begin{itemize} 
\item \texttt{Value},
\item  \texttt{Value\_U},
\item \texttt{Value\_Q}, 
\item \texttt{Value\_UU},
\item \texttt{Value\_UQ},
\item \texttt{Value\_QU},
\item \texttt{Value\_QQ}
\end{itemize}
 from \texttt{FunctionalInterface}, but also all the aforementioned methods with a preceeding \text{Point} (\texttt{PointValue} etc.).