### 2012-06-27 Examples/OPT/StatPDE/Example9/content.tex
Beispiel Optimierung mit FSI eingecheckt

In this example we solve an optimization problem where 
the equations come from fluid-structure interaction (FSI). The idea
is to extend the steady FSI benchmark problem (FSI 1) to an optimization 
problem where the drag is minimized over the cylinder plus beam.

The example has many similarities to Opt Example 3.

Detailed description will follow later.