### 2012-11-07 Paper_May_2012/main.tex
Habe versucht in den Seiten 6-8 etwas zu entzerren. Ausserdem gefallen mir die Paragraphen-Ueberschriften nicht, da die kursiv und sehr klein sind. Am liebsten wuerde ich die subsections aufwerten (Section 2.1: Solving of PDEs, Section 2.2: Solving of nonstationary PDEs, Section 2.3: solving of OPT. Dann hatten wir die subsection-Ueberschriften fuer localpde, etc. frei was momentan als Paragraph versteckt ist

\documentclass[prodmode,acmtoms]{acmsmall}
%

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

% Metadata Information
\acmVolume{0}
\acmNumber{0} 
\acmArticle{00}
\acmYear{0000}
\acmMonth{0}

%
%%% User-requested packages placed after this line %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsfonts,amssymb,amsmath}
\usepackage{booktabs,dcolumn}
\usepackage[T1]{fontenc}
\usepackage{mathtools}
\usepackage{enumerate,graphicx}
\usepackage{subfig}
\usepackage{multirow}
\usepackage{listings}

% \usepackage[center]{subfigure}
\numberwithin{equation}{section}
\usepackage{braket}                             %Provides \Set for typing sets
\usepackage{float} 
\usepackage{subfig}

%%% User's macros placed after this line %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\IfFileExists{dsfont.sty}%
%\usepackage{dsfont}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\renewcommand{\phi}{\varphi}
\renewcommand{\epsilon}{\varepsilon}
\newcommand{\deal}{\texttt{deal.II}}
\newcommand{\dope}{\texttt{DOpElib}}
\newcommand{\re}{\operatorname{Re}}

\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{black}{TODO: #1}}}}
%\newcommand{\mymarginpar}[1]{\marginpar{\textcolor{red}{\scriptsize{#1}}}}
\newcommand{\mymarginpar}[1]{\marginpar{\textcolor{red}{\normalsize{#1}}}}
%
\begin{document}
\markboth{C. Goll, T. Wick, and W. Wollner}{DOpElib: A Goal Oriented Software Library}

\title{DOpElib: A Goal Oriented Software Library for Computing PDEs and Optimization Problems}

\author{CHRISTIAN GOLL
\affil{University of Heidelberg}
THOMAS WICK
\affil{The University of Texas at Austin}
WINNIFRIED WOLLNER
\affil{University of Hamburg}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
In this article, we describe the software library 
\textit{Differential Equations and Optimization Environment} (\dope{}).
The main feature of \dope{} is that it provides a unified interface to high level algorithms 
such as time stepping methods, nonlinear solvers and optimization routines. This structure ensures 
that first of all the user is required to write those sections of code that are specific to 
the considered problem. Second, the exchange of parts of the used routines is possible 
with the need for just a few lines of code to change.
The article illustrates the performance and various features 
of the software package and provides some 
numerical results as demonstration.
\end{abstract}

%\category{C.2.2}{Computer-Communication Networks}{Network Protocols}

%\terms{Design, Algorithms, Performance}

%\keywords{Wireless sensor networks, media access control,
%multi-channel, radio interference, time synchronization}

%\acmformat{Zhou, G., Wu, Y., Yan, T., He, T., Huang, C., Stankovic,
%J. A., and Abdelzaher, T. F.  2010. A multifrequency MAC specially
%designed for  wireless sensor network applications.}

\begin{bottomstuff}
Author's adresses: C. Goll, Institut f\"ur Angewandte Mathematik,
Universit\"at Heidelberg;
T. Wick, The Institute for Computational Engineering and Sciences;
W. Wollner, Fachbereich Mathematik, Universit\"at Hamburg.
\end{bottomstuff}

\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{introduction}
Today, there exist a broad variety of software libraries
that provide numerical tools for solving partial differential
equations (PDEs). Prominent examples are 
\deal{} \cite{dealnew}, dune \cite{dune} , 
or commercial solvers like ANSYS \cite{ansys} or COMSOL \cite{comsol}. 
In fact, these libraries provide very useful tools 
for a huge variety of applications.

However, often, the user is confronted with 
certain bottlenecks. Either in open-source projects for academic research, 
standard high-level routines 
which are necessary to solve even simplest PDEs,
such as linear and nonlinear solvers or timestepping schemes must be implemented 
by the user himself. However, such routines are problem-independent 
and could be provided as standard tools. In fact, in many 
numerical studies, the main task is to compute a  
target functional which represents a certain error norm,
point value or traction, for example.
Second, there is nowadays an increasing interest in computing 
optimization problems in which the state equation is governed 
by a PDE. Such optimization problems comprise optimal control,
parameter estimation, inverse problems or optimal experimental design. 
So far, there exist a very limited number of open source packages
which provides standard routines that serve as basis for scientific studies.

On the other hand, 
commercial toolkits provide these previously mentioned routines but limits user-interaction
with the source code very often. This is quite unsatisfactory because
developement of novel numerical algorithms should enable the programmer
to change certain parameters or pieces of source code.

The novel software toolkit 
\textit{Differential Equations and Optimization Environment} (\dope{}) 
trys to fill this gap. We present a software
to solve forward PDE
problems as well as optimization problems constrained by a PDE. 
Its main feature is to give a unified interface to high level algorithms such as 
time stepping methods, nonlinear solvers and optimization routines. 
We aim that the user should only need to write those parts
of the code that are problem dependent while all invariant parts of the algorithms
should be reusable without any need for further coding.
In particular, the user should be able to switch between various different 
algorithms without the need to rewrite the problem dependent code, though he or she will
have to replace the algorithm object with an other one.

While \deal{} (which might presently seen as the basis of DOpE  
various finite elements are taken for the discretization) 
leaves much of the work and many
decisions to the user, \dope{} intends to be user-optimized by delivering
prefabricated tools which require from the user only adjustments connected
to his specific problem. The solution of optimization problems with PDE
constraints is an innovation in the \dope{} framework.
The focus is on the numerical solution of both stationary and nonstationary
problems which come from different application fields, like elasticity and
plasticity, fluid dynamics, and fluid-structure interactions.

The project was initially initiated at the University of Heidelberg in 2010 and  
at the present stage the following key features are supported by the library
\begin{itemize}
\item Solution of stationary and nonstationary PDEs in 1d, 2d, and 3d.
\item Various time stepping schemes (based on finite differences), 
  such as forward Euler, backward Euler,
  Crank-Nicolson, shifted Crank-Nicolson, and Fractional-Step-$\Theta$ scheme.
\item All finite elements of from \deal{} including hp-support.
\item Several examples showing the solution of several PDEs including
   Poisson, Navier-Stokes, Plasticity and fluid-structure interaction problems. 
\item Self written line search and trust region newton algorithms for the 
   solution of optimization problems with PDEs \cite{NoWr00}
\item Interface to SNOPT for the solution of optimization problems with PDEs and
  additional other constraints.
\item Several examples showing how to solve various kinds of optimization problems
  involving stationary PDE constraints.
\item Goal-oriented mesh adaptation with the Dual Weighted Residual method, see \cite{BR03}.
\item Different spatial triangulations for control and state variables.
\end{itemize}

The article is organized as follows. In the main Section
\ref{detailed_description}, a detailed description of 
the main features is given. In fact, we first explain the solution 
of stationary and nonstationary PDEs and give afterwards some 
ideas how to treat optimization problems. Finally, 
additional features are discussed. In the final Section
\ref{applications}, we discuss three different numerical 
examples which substantiates the features of \dope{}.

%The rest of this document is structured as follows: We start with an introduction in
%Chapter~\ref{chap:intro} where you will learn what is needed to run {\tt \dope{}}. 
%Further you will learn what problems we can solve and how all the different classes 
%work together for this purpose. This should help you figure out what the different classes
%do if you are in need of writing your own algorithm.
%
%Then assuming that you can work to your satisfaction with the algorithms already implemented
%we will show you how to create your own running example in Chapter~\ref{chap:howtoex}.
%This will be followed by a detailed description of all examples already shipped with 
%the library. You can find the examples for the solution of PDEs in Chapter~\ref{PDE}
%and those for the solution of optimization problems with PDEs in Chapter~\ref{OPT}.
%
%These notes conclude with a section that explains how we do automated testing of the 
%implementation in Chapter~\ref{chap:test}. This chapter will be of interest only if you 
%are trying to implement some new features to the library so that you can check that 
%the new code did not break anything.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Detailed Description of the Main Features}
\label{detailed_description}
%\todo{Aus der Sicht des Users anhand eines konkreten Beispiels beschreiben}
%\todo{Einfaches Beispiel nehmen}
%\todo{erwaehnen, dass wir alles immer nicht-linear betrachten}
%\todo{Aussehen eines typischen PDE und Opt loop und deren Gemeinsamkeiten
%erklaeren. Vielleicht mir Pseudocode}
This library is designed to allow easy implementation and numerical solutions 
of problems involving partial differential equations (PDEs). 
In particular, we remove all tedious routine implementation needed in 
\deal{} by introducing appropriate classes that take care of all the 
problem independent implementation while leaving only the problem specific 
part to the user. To show how this is designed we will, in the sequel, 
describe the reasoning first for a simple stationary PDE and then show
how this idea extends to non stationary problems as well as to optimization
problems with PDE constraints.

\subsection{Solving of PDEs}
\subsubsection{Stationary problems}\label{subsubsec:stationary problems}
For the start we consider a prototypical problem given by the solution of 
a stationary PDE, which we assume to be given in weak form.
In abstract notation it reads; find some $u$ in an appropriate set
such that
\begin{align}\label{eq:prototype_weak}
a(u)(\phi) = f(\phi) \quad \forall \phi \in V,
\end{align}
with some suitable space $V$ for the test functions and a 
right hand side $f$.
To illustrate this further, we take the stationary 
incompressible Navier-Stokes equations in a domain $\Omega \subset \R^d$, 
i.e., $u = (v,p)$ and $\phi = (\phi_u,\phi_p)$ then
\begin{align}\label{eq:ns}
a(u)(\phi) = \nu(\nabla v, \nabla\phi_u) + (v \nabla v,\phi_u) + (p, \nabla \cdot \phi_u) + (\nabla \cdot v ,\phi_p)
\end{align}
for all $\phi \in V = H^1_0(\Omega;\R^d) \times L^2_0(\Omega)$.
The solution will be searched for in $V + (u_d,0)$ where $u_d$ is some 
continuation of the given Dirichlet boundary data.
Then typically we do not only want to calculate $u$ but more likely we are 
interested in one (or several) functionals depending on the solution, i.e., some numbers $J(u)$.
%For example, if $u$ is the solution of the Navier-Stokes equations \eqref{eq:ns}, we look for the Drag 
%\begin{align*}
%J(u)=\int
%\end{align*}

In order to solve this problem, i.e., calculate $J(u)$, we need to discretize 
the equation. In \dope, we are using finite elements to do this. This means we need to pick a 
subdivision $\mathcal T_h$, for instance into quadrilaterals,
of the domain $\Omega$ and a finite element. In the concrete case we could, 
for instance, take the $\mathcal Q_2/\mathcal Q_1$ Taylor-Hood element.
As it is clear, that the initialization of the unknowns, the local shape 
functions, removal of hanging node constraints, etc. always follows the 
same procedure, we define an object which takes the finite element (FE)
and the triangulation and takes care of all other things related with 
handling the finite element space $V_h$. We will refer, to 
this as the \texttt{SpaceTimeHandler}. It is then clear, that the simplest 
form of the constructor should have the following interface:
\begin{lstlisting}
  template<typename Triangulation, typename FE>
      void SpaceTimeHandler(Triangulation &, FE);
\end{lstlisting}
Next, one needs to solve the discrete equation
\begin{align}\label{eq:discrete_equation}
a(u_h)(\phi_h) = f(\phi_h) \quad \forall \phi \in V_h.
\end{align}
To this end one can employ a newton type method with some globalization 
like line-search. To do so, one needs to calculate the residual of the 
equation, as well as solving linearized equations for the calculation of 
the newton method and then employing some kind of globalization strategy.
This again will usually require the evaluation of the residual.

The evaluation of the residual will require the calculation of the integrals 
involved in the weak formulation of the problem. For example 
the $i$-th component of the residual, with respect to the given FE-basis 
$\phi_{h,i}$, will read as follows
\begin{align}\label{eq:residual_vector}
 a(u_h)(\phi_{h,i}) - f(\phi_{h,i}) = \sum_{K\in \mathcal T_h} \nu\int_K \nabla u_h, \nabla \phi_{h,i}\,dx + \ldots
 \end{align}
Again, we see that, except for the local element integrals, all 
calculations are independent from the concrete form of the PDE. 
Hence we define an object \texttt{Integrator}
that offers a method to calculate the residual given a user-defined 
object \texttt{pde} that implements the local element integrals, i.e.,
\begin{lstlisting}
  template<typename PROBLEM,typename VECTOR>
      void Integrator
             ::ComputeNonlinearResidual(PROBLEM& pde, 
                                        VECTOR & residual);
\end{lstlisting}
where \texttt{VECTOR} is one of the possible vector types given by \deal{}.
It is clear that we need to ensure that the integrator object has access
to the information stored in the \texttt{SpaceTimeHandler}
to correctly loop over all elements.
To solve the linearized systems, again, all that is left is the calculation
of the matrix corresponding to the linearized equation. To calculate this,
 we can define a method in the integrator that does exactly this provided
we have a user defined object \texttt{pde} that can evaluate the local 
integrals needed for assembling the matrix, i.e.,
\begin{lstlisting}
  template<typename PROBLEM,typename MATRIX>
      void Integrator
             ::ComputeMatrix(PROBLEM& pde, 
                             MATRIX & system_matrix);
\end{lstlisting}
where now \texttt{MATRIX} is one of the matrix types supported by \deal{}.
With this we can now define a nonlinear solver \texttt{NewtonSolver} 
for the solution of the discrete problem requiring the existence of the 
following function
\begin{lstlisting}
 template<typename LINSOLVE>
  template<typename PROBLEM, typename VECTOR>
      void NewtonSolver<LINSOLVE>
             ::NonlinearSolve(PROBLEM& pde, 
                              VECTOR & solution);
\end{lstlisting}
Where \texttt{LINSOLVE} denotes the solver that should be used to solve the 
linearized equations during the newton iteration.

Finally, we note that to compute the requested functional value 
all that is left is to evaluate the functional in the computed solution 
which can be done again using an appropriate integrator method.
Since the sequence of events sketched above is again independent of the 
exact problem we introduce another container around it which we call 
\texttt{StatPDEProblem} that then simply provides the user with the 
method
\begin{lstlisting}
      void StatPDEProblem::ComputeReducedFunctionals();
\end{lstlisting}
That combines all the necessary steps to calculate $J(u_h)$.

Finally, we will need to have an object that is capable of providing all 
the necessary functionality required from the template argument 
\texttt{PROBLEM} in all of the above. We provide a wrapper named
\texttt{PDEProblemContainer} collecting all the problem specific information 
like the description of the PDE and the functional, the boundary values, etc. 
Although, having such an object instead of directly giving these informations
to the \texttt{StatPDEProblem} seems like a nuicance at the moment, we will 
see that it is of greater interest for time-dependent problems or 
optimization problems in what follows in Section~\ref{sec:timedep} and
Section~\ref{sec:opt} respectively.

\begin{remark}
In fact, the above constructors are not exactly what is implemented in \dope{}.
This is due to the fact, that for instance 
there are several different sparsity patterns available in 
\deal{} which the user should be able chose. Thus it needs to be 
given to the \texttt{SpaceTimeHandler}. However, to keep the 
presentation clear we refrained from including these details above and in
what follows.
\end{remark}

In the next section we will now sketch the way a user needs to implement the
user-dependent data as the local cell integrals and the initialization of the
objects described above.
\subsubsection{Problem dependent implementation of stationary PDE problems}
To solve a stationary PDE problem, the user has to provide some code which can roughly be divided into the following main sections:
\begin{description}
\item[Discretization] Specify what kind of triangulation,
% time steppingscheme,
finite element, quadrature rules, etc. we want to use.
\item[Problem-data] Define the weak form, the corresponding matrices, the functionals as well as the boundary conditions. With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
\item[Solvers] Which (non-)linear solver should we use? 
\item[Additional features] ParameterReader to read in the run-time parameters
  such as material coefficients or adjustments for the solvers like tolerance
  for Newton's method. Second, definition of Outputhandler etc.
\end{description}

We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force of the enclosed circle. We assume given Dirichlet data for the velocity $v$ on the boundary $\Gamma_{D}=\Gamma_{in}\cup\Gamma_{wall}\cup\Gamma_{circ}$. This is the Benchmark-Case 2D-1 from \cite{TuSchae96}, see there for further details.
%Additionally, for given inflow $v_{\operatorname{in}}$ we set  boundary conditions 
%v = v_{\operatorname{in}} \text{ on } \Gamma_{in}\text{ and }v = 0  \text{ on } \Gamma_{wall}.
\begin{figure}[h]
\centering
\resizebox{0.5\textwidth}{!}{\input{Pictures/example_ns}}
\caption{Flow around a cylinder with 
circle-center $C=(0.2,0.2)$ and radius $r=0.05$.}
\label{fig:example_ns}
\end{figure}

We split the program into the parts \texttt{main.cc}, \texttt{functionals.h} and \texttt{localpde.h}. The \texttt{main.cc}-file is responsible for the declaration of various objects which have mostly to do with the \textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form, the right hand side and the functional. We give a brief explanation on every file in the following.

\paragraph{\texttt{main.cc}}
In this file we gather the basic components that are necessary for the
solution process of a (stationary) partial differential equation. The
structure of the program is basically the same for every equation under
consideration, what changes is typically just the \textit{kind} of  finite
element or the linear solver. Here, a key issue is that we only need to pick 
some pre-implemented routines from the \dope{} library rather than 
implementing new features.

After the usual preamble, in which we include the necessary header-files from \deal{}, \dope{} and our own problem (i.e. \texttt{localfunctional.h} and \texttt{localpde.h}), we start by defining an instance of a \texttt{ParameterReader}-object, which handles run-time parameter.
\begin{lstlisting}
   ParameterReader pr;
\end{lstlisting}
In a text-file we might specify some problem-related parameters (for instance the viscosity $\nu$), but also general parameters for (non)-linear solvers and the output of computed data. 
After that, we create a triangulation of the domain $\Omega$ named \texttt{triangulation}.
%This is done exactly as in \deal{}, so we can choose between several possibilities: We could explicitly create a grid by providing a list of vertices and cells, read in a triangulation via \texttt{dealii::GridIn}-class or use the \texttt{dealii::GridGenerator}, which offers common domains like hypercubes or hyper-balls.

Next, we create some quadrature rules (for the evaluation of integrals over
cells and faces) and a finite element appropriate for the PDE at hand. We
choose a Gaussian quadrature of a certain degree and the previosly mentioned
Tayler-Hood element. These routines are quite standard in each finite element
software and we take advantage to take them from \deal{}.

%IntegratorDataContainer<VECTOR, 2> idc(quad_formula,
%                                         face_quad_formula);
%The \texttt{IntegratorDataContainer} class just makes the constructor interface of 
Then, the two characteristic features \texttt{LocalPDE} and \texttt{LocalFunctional} objects come into play:
\begin{lstlisting}

  LocalPDE<VECTOR> local_pde;
  LocalFunctional<VECTOR> local_functional;

\end{lstlisting}
These hold the information of the PDE and the functional under consideration. Further explication is given below because
the definition of this classes is done spererately in \texttt{localpde.h} resp.   \texttt{localfunctional.h}.

In the next step, we define an object of type \texttt{SpaceTimeHandler}, which handles all the things connected with the triangulation as well as the finite element space. This class is also capable of organizing degrees of freedom in time-dependent problems, hence the name. 
\begin{lstlisting}

  SpaceTimeHandler<TRIANGULATION, FE> dof_handler(triangulation,
                                                  finite_element);

\end{lstlisting}
There exist different versions of this class for forward as well as optimization problems.

Now we have everything at hand to create an instantiation of type \texttt{PDEProblemContainer}. This basically represents the disrectized problem, and bundles things like boundary conditions, the PDE, the functional and the discretization. First, we give the constructor the \texttt{local\_pde} and the \texttt{dof\_handler}.
\begin{lstlisting}

  PDEProblemContainer<LocalPDE> prob_container(local_pde,
                                               dof_handler);
\end{lstlisting}
After that, we add the functional under consideration and, because it is a boundaryfunctional, we specify the part of the boundary on which the functional operates (this is done via a 'boundarycolor', $1$ represents $\Gamma_{circ}$).
\begin{lstlisting}

  prob_container.AddFunctional(&local_functional);
  prob_container.SetBoundaryFunctionalColors(1);

\end{lstlisting}
Next, we incorporate the dirichlet data, which is given by the class \texttt{DirichleData}. Then we have to tell the \texttt{prob\_container} for which components and boundaries we prescribe the dirichlet conditions. This is done via a boundary-color (0 stands for $\Gamma_D$) and a \texttt{std::vector<bool> comp\_mask = (true, true, false)}.
\begin{lstlisting}

  DirichletData dd;
  prob_container.SetDirichletBoundaryColors(0, comp_mask, dd);

\end{lstlisting}
Now we create the object which steers the whole solution process, the previosly mentioned \texttt{StatPDEProblem}. It takes care of solving the discretized equation and evaluating the given functionals. The class depends actually on some template-parameters, so for the sake of abbreviation, we defined a typedef at the beginning of the \texttt{main.cc}-file.
\begin{lstlisting}

  typedef StatPDEProblem<NEWTONSOLVER,
           PDEProblemContainer<LocalPDE>, VECTOR, 2> SSOLVER;

\end{lstlisting}
Here, \texttt{NEWTONSOLVER} itself is an abbreviation of
\begin{lstlisting}

  typedef NewtonSolver<LINEARSOLVER, VECTOR, 2> NEWTONSOLVER;

\end{lstlisting}
The first template parameter of \texttt{NewtonSolver} describes which linear solver we use in this example. We opted for a direct solver, so: 
\begin{lstlisting}

  typedef DirectLinearSolverWithMatrix<MATRIX,
            VECTOR, 2> LINEARSOLVER;

\end{lstlisting}
With the type settled, we create the object \texttt{solver}, which needs the discretization (represented here by the \texttt{prob\_container}), the parameter reader as well as the quadrature rules.
\begin{lstlisting}

  SSOLVER solver(prob_container, pr,
                    quad_formula, face_quad_formula);

\end{lstlisting}
%Finally, the output handler is initialized via
%\begin{lstlisting}
%  P.RegisterOutputHandler(&out);
%  solver.RegisterOutputHandler(&out);
%\end{lstlisting}
Finally, the programm is started by
\begin{lstlisting}

  solver.ReInit();
  solver.ComputeReducedFunctionals();

\end{lstlisting}

\paragraph{\texttt{localpde.h}}
We shall now give an overview to the implementation of the specific form of the PDE.

%\marginpar{Rekapitulierung des Newton-Verfahrens?}
Let $N$ be the number of degrees of freedom of our discretization and $\Set{\phi_{h,i}|1\leq i \leq N}$ be a given FE-Basis. As stated in subsection~\ref{subsubsec:stationary problems}, to solve the discrete equation \eqref{eq:discrete_equation} we need the residual, i.e. the vector
\begin{align}\label{eq:discrete_residual}
\left(a(u_h)(\phi_{h,i})-(f,\phi_{h,i})\right)_{i=1}^N,
\end{align}
and the Jacobi-matrix of the weak form, i.e.
\begin{align}\label{eq:discrete_matrix}
\left(\nabla a(u_h)(\phi_{h,j},\phi_{h,i})\right)_{i,j=1}^N.
\end{align}
Now all we have to supply here are the cell-wise contributions to these terms, the \texttt{Integrator} takes care of assembling \eqref{eq:discrete_residual} and \eqref{eq:discrete_matrix}.

To this end, we define a class \texttt{LocalPDE} which provides the three methods
\texttt{CellEquation}, \texttt{CellRightHandSide} and \texttt{CellMatrix}. The implementation of the terms works by and large as in \deal{}. We show the basic structure of these methods using the example of \texttt{CellEquation}.

\begin{lstlisting}
  void CellEquation(const CellDataContainer<2>& cdc,
                    VECTOR &local_cell_vector)
    {
      const auto& fe_values = cdc.GetFEValuesState();
      unsigned int n_dofs_per_cell = cdc.GetNDoFsPerCell();
      unsigned int n_q_points = cdc.GetNQPoints();
\end{lstlisting}
 The \texttt{CellDataContainer} contains all the information we need to compute the cell-contributions (things like number of quadrature points, number of degrees of freedom on the cell, the \texttt{dealii::FEValues}-objects, etc.).
%  \begin{lstlisting}
%      _uvalues.resize(n_q_points,Vector<double>(3));
%      _ugrads.resize(n_q_points,vector<Tensor<1,2> >(3));
%      cdc.GetValuesState("last_newton_solution", _uvalues);
%      cdc.GetGradsState("last_newton_solution", _ugrads);
%  \end{lstlisting}
  \begin{lstlisting}
     cdc.GetValuesState("last_newton_solution", uvalues);
     cdc.GetGradsState("last_newton_solution", ugrads);
 \end{lstlisting}
 In these two lines, we load the (gradient of the) solution of the last newton iteration in all the quadrature points on the cell into the two vectors \texttt{uvalues} and \texttt{ugrads}.

In the next lines, we finally implement our concrete PDE. Here, aim to provide 
an interface which is closely related to the weak formulation which 
has to be derived by the user. Thus,
we loop over the quadrature points and the number of DoFs of the cell. In the innermost loop, we evaluate the weak form tested with $\phi_{i,h}$ (represented by \texttt{phi\_i\_grads\_v}) on the cell with the help of the previosly selected quadrature formula. 
%     const FEValuesExtractors::Vector velocities(0);
%     const FEValuesExtractors::Scalar pressure(2); 
 \begin{lstlisting}
    for(unsigned int q_point = 0; q_point < n_q_points; q_point++)
    {
      //initialize phi_i_grads_v etc.
      Tensor<2,2> vgrads;
      vgrads[0]= ugrads[q_point][0]; 
      vgrads[1]= ugrads[q_point][1];
      for (unsigned int i=0;i<n_dofs_per_cell;i++)
        local_cell_vector(i) += ( _nu 
                                * vgrads * phi_i_grads_v 
                                + //all the other terms)
                                * state_fe_values.JxW(q_point);
    }
\end{lstlisting}
\dope{} takes now care of the distribution of the local cell contributions into  the vector of the residual \eqref{eq:residual_vector}. The methods \texttt{CellRightHandSide} and \texttt{CellMatrix} follow the same structure as 
\texttt{CellEquation}, so we do not go into the details here.
% [CellEquation]
% 
% \begin{lstlisting}
%   void CellRightHandSide(const CellDataContainer<2>& cdc,
%                          dealii::Vector<double> &local_cell_vector,
%                          double scale);
% \end{lstlisting}
% 
% \begin{lstlisting}
%   void CellMatrix(const CellDataContainer<2>& cdc,
%                   dealii::FullMatrix<double> &local_entry_matrix);
% \end{lstlisting}
% The PDE under consideration is nonlinear, so let us briefly recapitulate manner such that for solving 
% the Laplace problem the following nonlinear step is necessary to recapitulate:
% Find $\delta u\in V$:
% \[
% (\nabla \delta u, \nabla \phi) = -(\nabla u, \nabla \phi) 
% \]
% and $u^{n+1} = u^n + \delta u$ for $n=1,2,3,ldots$.
% Therefore, the implementation reads for the right hand side:
% \begin{lstlisting}
%    void CellEquation(...)
%     {
%       // implement (\nabla u, \nabla \phi) 
%     }
% \end{lstlisting}
% Specifically, the inner loops reads
% \begin{lstlisting}
%    for (all quadrature points per cell q)
%       {
%         Tensor<2, 2> ugrads; // solution vector
%  
% 	for (all dofs per cell i)
%         {
% 	  const Tensor<2, 2> phi_i_grads_u; // test function
% 
%           local_cell_vector(i) += scale * 
%             scalar_product(ugrads, phi_i_grads_u) *
% 	    state_fe_values.JxW(q_point);
%         }
%       }
% \end{lstlisting}
% For the left hand side, we need to implement the Jacobian matrix:
% \begin{lstlisting}
%    void CellMatrix(...)
%     {
%       // implement  (\nabla \delta u, \nabla \phi)
%     }
% \end{lstlisting}
% Specifically,
% \begin{lstlisting}
%       for (all quadrature points per cell q)      
%         for (all dofs per cell i)    
%           for (all dofs per cell j)
%           {
%             local_entry_matrix(i, j) += 
%               scalar_product(phi_grads_u[j], phi_grads_u[i]) * 
%               state_fe_values.JxW(q_point);
%           }
% \end{lstlisting}
% The integration of the equation for each DoF and each quadrature point on each
% cell is then done 
% in the \texttt{templates/integrator.h} function. Finally, as usually in 
% finite element computations, this leads to a discrete system. Since
% we treat each problem with a Newton method, within each Newton step, the 
% following linear system must be solved:
% \[
% A\delta u = b
% \]
% in which $A$ is specified by \texttt{CellMatrix(...)} and 
% $b$ by \texttt{CellEquation(...)}. The vector $\delta u$ denotes the solution to the 
% present Newton step:
% \[
% u^n = u^{n-1} + \omega \delta u
% \]
% with a line search parameter $\omega\in (0,1]$.

\paragraph{\texttt{localfunctional.h}}
The implementation of the boundary-functional $J$ follows the outline of the computation of the cell-residual in \texttt{localpde.h}. However, there are two main differences. First, we do not have to specify the method \texttt{CellEquation}, but \texttt{BoundaryValue} instead. This is due to the fact that we consider a boundary-functional in this example, and so we have to specify the value of the functional not cell-wise but face wise (for the faces that lie on $\Gamma_{circ}$). The second difference is that as we only want to evaluate a functional and not a residual-vector, we just loop over the quadrature points and not over the degrees of freedom on the face.
\begin{lstlisting}
 double
 BoundaryValue(const FaceDataContainer<2>& fdc)
  {
    //same as before, get n_q_points etc.
    fdc.GetFaceValuesState("state", ufacevalues);
    fdc.GetFaceGradsState("state", ufacegrads);
\end{lstlisting}
Notice that we get here the values of the computed solution by giving the string "state".
\begin{lstlisting}
    for(unsigned int q_point = 0; q_point < n_q_points; q_point++)
      erg += -1. * ( _nu * ufacegrads[q_point][0]
                * state_fe_face_values.normal_vector(q_point)
                + //other terms)        
                * state_fe_face_values.JxW(q_point);
    return erg;
  }
\end{lstlisting}

% The target functional 
% is now implemented in \texttt{functional.h} and 
% treats a point evaluation:
% \begin{lstlisting}
%   double PointValue(...)
%    {
%     Point<2> p1(0.5, 0.5);
%     Vector<double> tmp_vector(2);
%     // more stuff is necessary here
%     VectorTools::point_value (...);
%     double x = tmp_vector(0);
% 
%     return x;
%    }
% \end{lstlisting}

% TWick:
% To solve the problem, we need a configuration (a mesh), eventually 
% a diffusion parameter, a finite element, a quadrature rule,
% a linear solver (here the CG solver), and some output such as
% a graphical solution and eventually some measurement functional.
% Each of those routines are written in the \texttt{main.cc} file. 
% We start by including necessary files both the \dope{} library and
% the \deal{} software:
% \begin{lstlisting}
%   #include "pdeproblemcontainer.h" etc. ...
% \end{lstlisting}
% In the \texttt{main} function, the following steps need to be done. First, 
% the parameter reader for runtime parameters might be initialized. In a parameter
% file the user might specify material parameters for the concrete PDE, but also
% general parameters for nonlinear and linear solvers, and output-related 
% aspects.
% 
% A linear finite element in two dimensions and two components is 
% initialized:
% \begin{lstlisting}
%   FESystem<2> state_fe(FE_Q<2>(1), 2);
% \end{lstlisting}
% and a quadrature rule of order three is used to approximate the 
% integrals:
% \begin{lstlisting}
%   QGauss<2> quadrature_formula(3);
% \end{lstlisting}
% Then, the \texttt{LocalPDE} object comes into play. This is an quite 
% important feature of \dope{} because it allows for easy implementation 
% of the variational formulation. Specific explication is given below because
% the implementation is separated in \texttt{localpde.h}.
% Afterwards, a \texttt{LocalPointFunctional} is declared. As for the PDE, the 
% implementation is done in another file, namely \texttt{localfunctional.h}.
% 
% Generically, each problem in \dope{} is considered to be time-dependent and nonlinear because
% these are the 'real' world problems. Therefore, (even for the linear and stationary Laplace equation),
% we have to declare a (pseudo) time:
% \begin{lstlisting}
%   std::vector<double> times(1, 0.);
% \end{lstlisting}
% Then, the spatial mesh (here the unit square) is created and three-times globally-refined:
% \begin{lstlisting}
%   GridGenerator::hyper_cube(triangulation, 0, 1);
%   triangulation.refine_global(3);
% \end{lstlisting}
% A major component which organizes the whole solution process (independent from 
% the problem!) is the 
% \begin{lstlisting}
%   MethodOfLines_StateSpaceTimeHandler<...> dof_handler (...);
% \end{lstlisting}
% This method handles all dofs in space and time and different versions exist
% for forward as well as optimization problems.
% 
% Afterwards, the problem (here the Laplace PDE) is given to the \texttt{dof_handler} via
% \begin{lstlisting}
%   OP P(local_pde, dof_handler);
% \end{lstlisting}
% It remains to declare boundary conditions for the problem under consideration. 
% Let's simply take zero Dirichlet conditions on the boundary:
% \begin{lstlisting}
%   std::vector<bool> comp_mask(2);
% 
%   comp_mask[0] = true;
%   comp_mask[1] = true;
% 
%   DOpEWrapper::ZeroFunction<2> zf(2);
%   SimpleDirichletData<VECTOR, 2> DD1(zf);
% 
%   P.SetDirichletBoundaryColors(0, comp_mask, &DD1);
% \end{lstlisting}
% Next, the solver for the linear equations needs to be specified:
% \begin{lstlisting}
%   SSolver solver(&P, "fullmem", pr, idc);
% \end{lstlisting}
% Finally, the output handler is initialized via
% \begin{lstlisting}
%   P.RegisterOutputHandler(&out);
%   solver.RegisterOutputHandler(&out);
% \end{lstlisting}
% Finally, the programm is started with:
% \begin{lstlisting}
%   solver.ReInit();
%   out.ReInit();
%   out.Write(outp, 1, 1, 1);
%   solver.ComputeReducedFunctionals();
% \end{lstlisting}
% After explaining the basic compononents that are necessary (and always the same in \dope{}) 
% for all  (linear and stationary) partial differential equations, we shall give 
% an overview to the implementation of the specific form of the PDE. 
% Beside specification of boundary conditions, the form of the local PDE depends 
% on the concrete problem under consideration. 
% This step is done in the \texttt{localpde.h} file. As already explained,
% each PDE is treatet in a nonlinear manner such that for solving 
% the Laplace problem the following nonlinear step is necessary to recapitulate:
% Find $\delta u\in V$:
% \[
% (\nabla \delta u, \nabla \phi) = -(\nabla u, \nabla \phi) 
% \]
% and $u^{n+1} = u^n + \delta u$ for $n=1,2,3,ldots$.
% Therefore, the implementation reads for the right hand side:
% \begin{lstlisting}
%    void CellEquation(...)
%     {
%       // implement (\nabla u, \nabla \phi) 
%     }
% \end{lstlisting}
% Specifically, the inner loops reads
% \begin{lstlisting}
%    for (all quadrature points per cell q)
%       {
%         Tensor<2, 2> ugrads; // solution vector
%  
% 	for (all dofs per cell i)
%         {
% 	  const Tensor<2, 2> phi_i_grads_u; // test function
% 
%           local_cell_vector(i) += scale * 
%             scalar_product(ugrads, phi_i_grads_u) *
% 	    state_fe_values.JxW(q_point);
%         }
%       }
% \end{lstlisting}
% For the left hand side, we need to implement the Jacobian matrix:
% \begin{lstlisting}
%    void CellMatrix(...)
%     {
%       // implement  (\nabla \delta u, \nabla \phi)
%     }
% \end{lstlisting}
% Specifically,
% \begin{lstlisting}
%       for (all quadrature points per cell q)      
%         for (all dofs per cell i)    
%           for (all dofs per cell j)
%           {
%             local_entry_matrix(i, j) += 
%               scalar_product(phi_grads_u[j], phi_grads_u[i]) * 
%               state_fe_values.JxW(q_point);
%           }
% \end{lstlisting}
% The integration of the equation for each DoF and each quadrature point on each
% cell is then done 
% in the \texttt{templates/integrator.h} function. Finally, as usually in 
% finite element computations, this leads to a discrete system. Since
% we treat each problem with a Newton method, within each Newton step, the 
% following linear system must be solved:
% \[
% A\delta u = b
% \]
% in which $A$ is specified by \texttt{CellMatrix(...)} and 
% $b$ by \texttt{CellEquation(...)}. The vector $\delta u$ denotes the solution to the 
% present Newton step:
% \[
% u^n = u^{n-1} + \omega \delta u
% \]
% with a line search parameter $\omega\in (0,1]$.
% 
% 
% The target functional 
% is now implemented in \texttt{functional.h} and 
% treats a point evaluation:
% \begin{lstlisting}
%   double PointValue(...)
%    {
%     Point<2> p1(0.5, 0.5);
%     Vector<double> tmp_vector(2);
%     // more stuff is necessary here
%     VectorTools::point_value (...);
%     double x = tmp_vector(0);
% 
%     return x;
%    }
% \end{lstlisting}

\subsubsection{Nonstationary problems}\label{sec:timedep}
To describe the solution of nonstationary 
PDEs, we consider a time-dependent semi-linear form $a(u)(\phi)$ and some
right hand side functional $l(\phi)$.
Let $I:=[0,T]$ be a time interval with end time point value $T$.
To find some $u$ in a weak form, we define
\[
\int_I a(u)(\phi)\, dt = \int_I l(\phi)\, dt \quad \forall\Phi\in X,
\]
with some suitable Bochner space $X$.  As model problem,
we consider the nonstationary Navier-Stokes equations
in a domain $\Omega\in\mathbb{R}^d$. Find: 
$(v,p)\in X$ with
\[
\int_I \bigl[ (\partial_t v, \phi)
+ \nu (\nabla v, \nabla \phi) + (v\cdot\nabla v, \phi)
- (p,\nabla\cdot \phi)
+ (\nabla\cdot v, \chi)\bigr] \, dt
+ (v(0) - v^0, \phi(0))
= \int_I l(\phi) \, dt 
\]
for all $\Phi:= (\phi, \chi) \in X$ with
\[
X:= \bigl\{ 
(v,p) | v\in L^2(I,H_0^1(\Omega)^d) , 
\partial_t v\in L^2(I, L^2(\Omega)^d), 
p\in L^2(I,L^2(\Omega)\setminus\mathbb{R} \bigr\}.
\]
As for stationary problems, the goal is to compute
some target functional $J(\cdot)$ which is now possibly 
time-dependent. To organize the discretization 
process, we need to perform spatial and temporal 
discretization. This is again done via
an object that takes care of organizing the degrees of 
freedom, as in the case of stationary problems.
The only difference to the stationary case is that we need to 
add the temporal subdivision \texttt{times}. In the simplest 
case, when we have the same spatial triangulation 
at all times, the constructor could look as follows:
\begin{lstlisting}
  template<typename Triangulation, typename FE, typename times>
      void SpaceTimeHandler(Triangulation &, FE, times&);
\end{lstlisting}

Now we may discretize the problem with some time-stepping scheme, which will 
remove the temporal derivative and replace it by some difference quotients. 
Since the exact way of doing this depends on the time-stepping scheme 
we need a new object that takes the user-given description of the PDE and 
generates the appropriate combinations of spatial integrals and scaling.
Thus a generic problem may look as follows:
\begin{lstlisting}
  template<typename PROBLEM>
      class TimeStepProblem;
\end{lstlisting}
At present we have prepared the following time stepping schemes in \dope:
\begin{itemize}
\item Forward-(Explicit)-Euler
\item Backward-(Implicit)-Euler
\item Crank-Nicolson scheme
\item shifted Crank-Nicolson scheme (to deal with singular initial data)
\item fractional-step-$\theta$ scheme
\end{itemize}
With this we can now run a loop over all timesteps, where in each of the 
timesteps we can reuse the functionality to solve stationary PDEs 
applied to the time-step problem.

Now it is apparent, why we introduced the \texttt{StatPDEProblem}, because
we now define the \texttt{InstatPDEProblem} to contain the necessary 
loop over all time points. Next, we will describe the methods the user needs to 
provide in addition to the stationary case.

\subsubsection{Problem-dependent 
implementation of nonstationary problems}
\label{sec:timedep:implementation}
We explain in more detail the temporal discretization
and specific issues of its implementation. 
To keep the presentation easy, we abuse notation 
by neglecting
all material coefficients
and give formal-related idea based on the weak 
form of the Navier-Stokes equations:
Find $(v,p)$ such that
\begin{align*}
(\partial_t v,\phi) 
+ (\nabla v, \nabla \phi)
+ (v\cdot\nabla v,\phi)
-(p,\nabla\cdot \phi)
+(\nabla\cdot v, \chi)
=(f,\phi),
\end{align*}
where $f$ describes some volume forces.
Temporal discretization using the One-step-$\theta$ scheme reads:
Given the old timestep solution $(v^n,p^n)$, 
find $(v^{n+1}, p^{n+1})$ (and multiplication 
by the inverse of the timestep $k$):
\begin{align*}
(v^{n+1} - v^{n}, \phi)
&+ k\theta (\nabla v^{n+1}, \nabla \phi)
+ k\theta (v^{n+1}\cdot\nabla v^n + 
  v^{n}\cdot\nabla v^{n+1},\phi)\\
&- k (p^{n+1},\nabla\cdot \phi)
+ (\nabla\cdot v^{n+1}, \chi)\\
=&\; k\theta (f^{n+1},\phi) + k(1-\theta) (f^{n},\phi)
- k(1-\theta) (\nabla v^{n}, \nabla \phi) 
\end{align*}
The concrete choice of $\theta$ leads to a
specific time stepping scheme. For instance,
$\theta = 0.5$ gives the Crank-Nicolson scheme,
\begin{lstlisting}
   CrankNicolsonProblem<...>
\end{lstlisting}

The pressure and incompressibily terms should be 
treated fully implicitly when time-discretizing 
the Navier-Stokes equations.
For this reason, the implementation is split up
into several items to account for fully implicit 
terms and remaining terms for which we need 
old timestep and present timestep values. 
Hence, in the local cell equation, the parameters
\texttt{scale} and \texttt{$scale_ico$}
are used to distigugish these terms:
\begin{lstlisting}
  void CellEquation(...)
   {
      local_cell_vector(i) +=  
          scale * (\nabla v^{n+1}, \nabla \phi) + etc.

      local_cell_vector(i) +=  
          scale_ico * (-p^{n+1},\nabla\cdot v)
   }
\end{lstlisting}
Analogoues procedure is performed in the local 
cell matrix.

In addition to stationary problems, the implementation
of the time-derivative must be considered which is done
in 
\begin{lstlisting}
  void CellTimeEquation(...)
   {
     local_cell_vector(i) += 
         scale * (u^{n+1},\phi) + etc. 
   }
\end{lstlisting}
and it is done in a straight-forward way:
\begin{lstlisting}
  for (all quadrature points per cell q)
   {
     Tensor<1,2> u = ...;

     for (all dofs per cell i)
      {
        const Tensor<1,2> phi_i_u = ...;

        local_cell_vector(i) +=  
          scale * (u * phi_i_u) * 
          state_fe_values.JxW(q_point);
      }
   }
\end{lstlisting}
Here, the advantage is that \dope{} orders the terms in the timestepping
problems
such that only $(u^{n+1},\phi)$ must be implemented.

\begin{remark}[Nonlinear terms in the time derivative]
However, if someone 
needs explicit input of the old timestep term, then he can use 
\begin{lstlisting}
  void CellTimeEquationExplicit (...)
   {
     // implement (u^{n+1},\phi) and (u^n,\phi)  
   }
\end{lstlisting}
\end{remark}

\begin{remark}[Fractional-Step-$\theta$-scheme]
The temporal discretization in this section 
is based on the One-Step-$\theta$-schemes. The library
offers in addition an implementation 
of the second-order and strongly $A$-stable 
Fractional-Step-$\theta$-scheme in which 
each time step is subdivided into three 
sub-timeintervals. 
\end{remark}

\subsection{Solving of optimization problems}\label{sec:opt}
Upgrading the framework for solving optimization problems
is described in this section. A prototypical problem reads:
\begin{align*}
\min\;&J(q,u) \\
  &\text{s.t.}\; a(q,u)(\phi) = 0 \quad \forall \phi\in V,\\
  &a \le q \le b,\\
  &g(q,u) \le 0,  
\end{align*}
where $u$ is a FE-function and $q$ can either be a FE-function or some 
fixed number of parameters, $a$ and $b$ are constraint bounds for the control $q$,
and $g(\cdot)$ is some state constraint.
To keep the presentation brief we will neglect the additional bounds
\[
a \le q \le b,\qquad g(q,u) \le 0
\]
although they can be treated by \dope{} as well.
Then, in order to minimize the functional $J$ one may apply the, so-called, 
reduced approach, in which we eliminate the dependent variable $u$ by the solution 
operator $S$ of the PDE, i.e., 
\begin{align*}
a(q,S(q))(\phi) = 0 \quad \forall \phi\in V. 
\end{align*}
This leads to the problem
\begin{align*}
\min\;&j(q) = J(q,S(q)). 
\end{align*}
Then, one can apply any algorithm to solve the, unconstraint, minimization 
problem. This, will in any case require the evaluation of $j(q)$ which 
we can do, e.g., using \texttt{StatPDEProblem::ComputeReducedFunctional} for 
a stationary PDE. If one would like to use derivative based methods, one 
needs in adition derivatives of $j$ and thus of $J$ as well as of $S$. 
To do so, we remark, that one can calculate the derivative of 
$j$ using the adjoint calculus. Then the directional derivative 
of $j$ in direction $\delta q$ is given by
\[
j'(q)(\delta q) = J'_q(q,S(q))(\delta q) + (S^* J_u'(q,S(q)),\delta q)
\]
where the index $q$ or $u$ denote directional derivatieves for the 
respective argument. To calculate action of the adjoint operator $S^*$ 
we note that $z =  S^* J_u'(q,S(q))$
can be computed via the adjoint PDE
\begin{equation}
  \label{opt_KKT_system}
  \begin{aligned}
    a_{u}'(q,u)(\phi , z) &= J'_{u}(q,u)(\phi) & \forall \phi&\in {\cal X}.
  \end{aligned}
\end{equation}
Similarly formulas for the second derivatives can be obtained.

To do so, we define an object 
\begin{lstlisting} 
  template<typename PDE, typename COSTFUNCTIONAL>
    class OPTProblemContainer;
\end{lstlisting}
that collects the user given PDE and functional descriptions and constructs the 
required PDE and right hand side information needed for the calculations.

This means for the user only the derivatives of the functional and the PDE 
needs to be provided in addition to the 
\begin{lstlisting} 
  void CellEquation_U(...)
   {
     // implement directional derivative of the CellEquation
   }
\end{lstlisting}
and functional terms, i.e.,     
\begin{lstlisting} 
  virtual void BoundaryValue_U(...)
      {
        // implement here the derivative of 
        // the boundary cost functional
      }
\end{lstlisting}

Due to the design of the classes it is easily possible to extend any given 
and running code for a PDE both to the nonstationary counterpart, if the PDE is 
stationary, as well as to an optimization problem where the constraint is 
given by this PDE. Meaning that the use of our framework for the simulation of 
PDEs will, due to the enforced interface, enable easy application of state-of-the-art 
optimization algorithms to minimization problems governed by the PDE.

\subsection{Additional features}
Finally, we shall give a brief summary of features 
that are applicable (in general) to all types of problems. Specifically,
\dope{} provides goal-oriented mesh refinement with the help of the 
Dual Weighted Residual (DWR) method, see \cite{BeRa96} or \cite{BR03}. Here, a certain 
cost functional such as a point evaluation, line intergration or domain
integration is considered. Determining the lowest error between the 
(unknown) exact solution and its numerical approximation for lowest computational
cost is the main goal of the DWR method. The same routines 
can also be used, without any modification, to use standard residual error estimators by exchanging the 
error estimator class.

Since in many applications it is desirable to use different discretizations
for the state and the control \dope{} provides multi-mesh support, meaning that
onec a problem has been written correctly using the same mesh for all finite element
variables one can switch to a different SpaceTimeHandler that allows different meshes for all 
variables without the need to change the user-given implementation of the PDE and functional.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Applications}
\label{applications}
In this final section, 
we present three numerical tests which demonstrate different
features of \dope{}:
\begin{itemize}
\item In the first numerical example, we consider 
goal-oriented mesh refinement with the help of the 
DWR method for the stationary Navier-Stokes equations.
\item The second example presents nonstationary fluid-structure 
interaction. The challenges for those kind of problems are the multi-domain
character of the coupled problem and the treatment of the coupling conditions.
The problem is formulated in a variational monolithically-coupled way which 
allows to consider goal-oriented mesh refinement and gradient-based optimization.
\item In the third numerical test, an optmization problem for structural mechanics
is discussed.
\end{itemize}

\subsection{Goal-oriented mesh refinement for Navier-Stokes}
In this example we consider a laminar flow around a cylinder in 2d. To be more precise, we are computing the benchmark 2D-1 from \cite{TuSchae96}. The computational domain is as depicted in figure~\ref{fig:example_ns}. We want to compute the drag force on the enclosed cylinder
\begin{equation}
c_D = \frac 1 {20} \int_{\Gamma_{circ}} \nu\partial_nu _1 - pn,
\end{equation}
with the normal $n$ and $\nu = 0.001$. To achieve an efficient computation of the quantity $c_D$, we employ goal-oriented mesh refinement with the help of the DWR-method, see \cite{BeRa96}.

\begin{itemize}
\item Bild Freiheitsgrade vs Fehler, global und 
\item Bild adaptive mesh, bild der Loesung (?)
\item eventuell kurz auf duale gleichung eingehen.
\end{itemize}

\subsection{Nonstationary Fluid-Structure Interaction}
In this example, a challenging benchmark FSI 2
proposed by Hron and Turek \cite{HrTu06b} is considered.
To solve this problem accurately, it is very important that 
the coupling conditions
\[
v_f = v_s \quad \text{and} \quad \sigma_f n_f = \sigma_s n_s, 
\]
namely, the continuity of velocities and continuity of normal stresses
are satisfied in each time step. This is achieved with the help of 
a monolithic coupling scheme (strong coupling).

The basic configuration is 
sketched in Figure \ref{fig:example_ns} at which an elastic beam is attached 
behind the rigid cylinder. 
The computational domain has length $L=2.5m$ and height $H=0.41m$. The circle center
is positioned at $C=(0.2m,0.2m)$ with radius $r=0.05m$. The elastic beam has length
$l=0.35m$ and height $h=0.02m$. The right lower end is positioned at 
$(0.6m,0.19m)$, and
the left end is attached to the circle. 
%\begin{figure}[h]
%\centering
%\input{Pictures/pic_12}
%\caption{Flow around cylinder with elastic beam with 
%circle-center $C=(0.2,0.2)$ and radius $r=0.05$.}
%\label{configuration_csm_and_fsi_2D}
%\end{figure}
Control points $A(t)$ (with $A(0) = (0.6,0.2)$) are fixed at the 
trailing edge of the structure, measuring $x$- and $y$-deflections of the beam.

Details 
on parameters and evaluation functionals and other results 
can be found in \cite{HrTu06b,BuSc06,DeHaeAnnBrVie10,Wi11}. 
The Fractional-Step-$\theta$ scheme is used for time discretization with
different time step sizes $k$. However, the time stepping scheme can be 
very easily chosen in the \texttt{main.cc} function by choosing an appropriate 
time stepping scheme, at present the following are availiable.
\begin{lstlisting}
ForwardEulerProblem
BackwardEulerProblem
CrankNicolsonProblem
ShiftedCrankNicolsonProblem
FractionalStepThetaProblem
\end{lstlisting}

The quantities of interest are evaluations of 
$x$- and $y$ displacement at the point $A(0) = (0.6,0.2)$
and the drag and lift forces acting on the cylinder and the elastic beam:
\begin{align}
\label{drag_lift_forces}
(F_D , F_L) 
= {\int_{S_f} \sigma_f \cdot n_f \, ds + 
\int_{\Gamma_i} \sigma_s \cdot n_s \, ds},
\end{align}
where $S_{f}$ denotes the path over the cylinder in the fluid part and
$\Gamma_i$ the interface between the elastic beam and the 
fluid.

Since the problem is fully nonstationary, the 
dynamics are presents in Figure \ref{res:fsi_2_mesh_and_x_velo}. 
By comparison of our findings in Figure \ref{res:results_ux_and_uy_fsi_2}
with the literature, it can be identified that 
the implementation in \dope{} is correct.

% Bilder von FSI 2 
\begin{figure}[h]
\centering
{\includegraphics[width=6cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_mesh8070_scale.png}}
{\includegraphics[width=6cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_x_velo8070_scale.png}}
\caption{FSI 2 test case: mesh (left) and velocity profile in vertical 
direction (right) at time $t=16.14s$.}
\label{res:fsi_2_mesh_and_x_velo}
\end{figure}

\begin{figure}
\centering
{\includegraphics[width=5.8cm]{Pictures/ux_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
{\includegraphics[width=5.8cm]{Pictures/uy_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
{\includegraphics[width=5.8cm]{Pictures/Drag_fluid_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
{\includegraphics[width=5.8cm]{Pictures/Lift_fluid_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
\caption{FSI 2: the deflections of the beam, $u_x(A)$ and $u_y(A)$ (in $cm$), and 
the drag $F_D$ and the lift $F_L$ evaluation (in $kg/m\,s^2$) are displayed versus
time (in $s$).
} 
\label{res:results_ux_and_uy_fsi_2}
\end{figure}

\newpage
\subsection{Compliance minimization}
In this example, we consider the compliance minimization of a standard MBB-Beam. 
The thickness variable is allowed to take intermediate value apart from $0$ and $1$,
i.e., we consider a variable thickness sheet. The discretization is done 
using $Q_2$ elements for the displacement and discontinuous $P_0$ elements for the 
thickness.
\begin{figure}
\centering
\includegraphics[width=1.\textwidth, viewport=150 0 1024 500, clip]{Pictures/MBB.png}
\caption{Thickness distribution for the MBB-beam minimum compliance problem.} 
\label{res:mbb}
\end{figure}
A typical thickness distribution is shown in Figure~\ref{res:mbb}.

\section*{Thanks}
The \dope{} project makes use of various finite elements taken from  
the \deal{} \cite{dealnew} finite element library which has been developed
 initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{dealold}.
The authors acknowledge their past experience as well as discussions 
on modularization of algorithms
with 
the authors of the libraries 
Gascoigne/RoDoBo project, which was initiated by 
Roland Becker, Dominik Meidner, and Boris Vexler \cite{rodobo}. 
The second author thanks Mary F. Wheeler (ICES, Austin) for the 
possibility to finish this work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\bibliographystyle{abbrvnat}
\bibliographystyle{acmsmall}
\bibliography{lit}
%

\end{document}