### 2012-12-05 Paper_May_2012/LetterToTheEditior.tex

\documentclass{uhhbrief-fak6}    % Deutsch

\usepackage[utf8]{inputenc}

\usepackage{graphicx}
\usepackage{amsthm,amsmath,amsfonts,amssymb}
\usepackage{watermark}
\usepackage{url}

\selectlanguage{english} %For english Language

\name{Prof. Dr. Winnifried Wollner}
\signature{Winnifried Wollner}
\fbname{Fachbereich Mathematik}
\fbabkuerzung{FB Mathematik}
\strasse{Bundestr. 55}
\plzort{20146 Hamburg}
\telefonnummer{+49 (0)40 42838}
\telefondurchwahl{5988}
\faxnummer{+49 (0)40 42838}
\faxdurchwahl{5117}
\email{winnifried.wollner@math.uni-hamburg.de}
\http{www.math.uni-hamburg.de/home/wollner}

\date{\today}

\begin{document}
\thiswatermark{\begin{minipage}{\textwidth}\vspace*{43.5cm}\hspace*{5.8cm}\includegraphics[width=0.8\textwidth]{BG_logo}
  \end{minipage}
}
\begin{letter}{Editor-in-Chief of Computing\\Prof. Dr. Schahram Dustdar\\[\baselineskip]
    electronic}

\subject{Paper Submission}

\opening{Dear Professor Dustdar,}

please find enclosed the manuscript entitled

{\em DOpElib: Differential Equations and Optimization Environment; A Goal Oriented Software Library for Solving PDEs and Optimization Problems with PDEs}

for publication in Computing by
C.Goll, T. Wick and W. Wollner.

The article outlines the guiding design principles of our software environment
DOpElib (see \url{dopelib.net}) for the numerical solution of PDEs and optimization problems subject 
to PDEs with special emphasis on modular design of all components. 
We hope that you deem this publication worthy of publication in Computing.

\closing{Yours sincerely,}

%  \vspace*{-25mm}\includegraphics[width=.30\textwidth]{Signatur}\\[-2cm] $\mbox{}$\vspace{24mm}

%\ps{some ps}
%\cc{Some CC adresses}

%\encl{Some Anlagen}

\end{letter}
\end{document}