### 2013-01-10 Examples/PDE/StatPDE/Example11/content.tex
Forgot at last checking

\subsubsection{General problem description}
\index{stationary PDE}
In this example we consider the same setting as in subsection~\ref{PDE_Stat_Stokes}, the only
difference is that we want to employ the hp-Finite-Elements. So the equation we solve is still 
the stationary incompressible Stokes equation \index{stationary PDE!Stokes equation}. Here,
we use the symmetric stress tensor which has a little consequence when using 
the do-nothing outflow condition. In strong formulation we have
\begin{align} \label{eq:strong}
-\frac{1}{2}\nabla\cdot (\nabla v + \nabla v^{T}) + \nabla p &= f \\ 
\nabla \cdot v &= 0 \notag
\end{align} 
on the domain $\Omega = [-6,6]\times [0,2]$. We choose for simplicity $f=0$.\\

\subsubsection{Adding hp-Elements}
One sees by comparing  the \texttt{main.cc}-file of this problem with the one of subsection~\ref{PDE_Stat_Stokes} that the change to hp-Elements is really easy, so we will keep the description short.

The obvious things to change are some templa