### 2013-04-10 Examples/OPT/StatPDE/Example1/content.tex
Added DWR-Estimator for Optimization Problems (CostFunctional only, others will come in time...). See OPT/StatPDE/Example1 for details.

\subsubsection{General problem description}
This example solves the distributed minimization problem
\begin{gather*}
\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
\text{s.t.} (\nabla u,\nabla \phi) = (q+f,\phi)\;\;\forall\,\phi \in H^1_0(\Omega)
\end{gather*}
on the domain $\Omega = [0,1]^2$, and the data is chosen as follows:
\begin{gather*}
 f = \left(20\pi^2  \sin(4 \pi x) - \frac{1}{\alpha}  \sin(\pi x)\right) \sin(2 \pi y)\\
 u^d = \left( 5 \pi^2 \sin(\pi x) + \sin(4 \pi x)\right)  \sin(2\pi y)
\end{gather*}
and $\alpha = 10^{-3}$.
Hence its solution is given by:
\begin{gather*}
 \overline{q} = \frac{1}{\alpha} \sin(\pi x) \sin(2 \pi y)\\
 \overline{u} = \sin(4 \pi x) \sin(2 \pi y).
\end{gather*}
Thus the exact optimal value of the cost functional can be calculated as 
\begin{gather*}
 J^* = J(\overline{q},\overline{u}) = \frac{1}{8}\Bigl(25\pi^4 + \frac{1}{\alpha}\Bigr).
\end{gather*}

In addition the following functionals are evaluated:
\begin{gather*}
  \text{MidPoint: } u(0.5 ; 0.5)\\[2mm]
  \text{MeanValue: }\int_\Omega u
\end{gather*}

The example shows how to estimate the error in the cost functional for stationary optimization
problems.