### 2013-05-08 Examples/OPT/StatPDE/Example7/content.tex
Added support for IPOPT

\subsubsection{General problem description}
This example solves the distributed minimization problem
\begin{gather*}
\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
\text{s.t.} (\nabla u,\nabla \phi) = (q+f,\phi)\;\;\forall\,\phi \in H^1_0(\Omega)\\
\text{s.t.} -500 \le q \le 500\;\text{a.e. in }\Omega
\end{gather*}
on the domain $\Omega = [0,1]^2$, and the data is chosen as follows:
\begin{gather*}
 f = \left(20\pi^2  \sin(4 \pi x) - \frac{1}{\alpha}  \sin(\pi x)\right) \sin(2 \pi y)\\
 u^d = \left( 5 \pi^2 \sin(\pi x) + \sin(4 \pi x)\right)  \sin(2\pi y)
\end{gather*}
and $\alpha = 10^{-3}$.

In addition the following functionals are evaluated:
\begin{gather*}
  \text{MidPoint: } u(0.5 ; 0.5)\\[2mm]
  \text{MeanValue: }\int_\Omega u
\end{gather*}

The Problem is similar to that of {\tt OPT/StatPDE/Example1} except for the box control constraints.
The problem is solved using the optimization library IPOPT that you can obtain for free. To use it 
a correct link to the ipopt library needs to be created in  {\tt DOpE/ThirdPartyLibs} by the name 
{\tt ipopt}, i.e., you should have the file
{\tt DOpE/ThirdPartyLibs/ipopt} pointing to the ipopt directory. If you have not done this you can compile the
example but when running the example you will only get an error message like\\
{\tt Warning: During execution of `Reduced\underline{ }IpoptAlgorithm::Solve`\\ the following Problem occurred!\\
To use this algorithm you need to have IPOPT installed! \\To use this set the WITH\underline{ }IPOPT CompilerFlag.}

Alternatively the commercial optimization library SNOPT can be used in this example.
In order to use this library you need to install SNOPT on your computer and then generate a symlink to 
the snopt directory (where you have the libs and the header files) 
in the {\tt DOpE/ThirdPartyLibs} directory named {\tt snopt}, i.e., you should have the file
{\tt DOpE/ThirdPartyLibs/snopt} pointing to the snopt directory. If you have not done this you can compile the
example but when running the example you will only get an error message like\\
{\tt Warning: During execution of `Reduced\underline{ }SnoptAlgorithm::Solve`\\ the following Problem occurred!\\
To use this algorithm you need to have SNOPT installed! \\To use this set the WITH\underline{ }SNOPT CompilerFlag.}