### 2013-05-09 Examples/PDE/StatPDE/Example7/content.tex
Further work on the examples

\subsubsection{General problem description}
This is the simple vector valued laplace equation on the unitsquare with righthandside $(1,1)$ and zero dirichlet values. The classical example of a PDE.