### 2013-05-10 Examples/PDE/StatPDE/Example7/content.tex
Further work on the examples

\subsubsection{General problem description}
In this problem we solve the simple vector valued laplace equation in 2d
on the unit square $\Omega=[0,1]^2$, i.e. in strong formulation we look for $u=(u_1,u_2)$ s.t.
\begin{align*}
-\Delta u =& f &&\text{in }\Omega.
\end{align*}
We set zero dirichlet values on $\partial \Gamma$ and choose $f=(1,1)$. The classical example of a PDE.