### 2013-05-10 Examples/OPT/StatPDE/Example3/content.tex
Changed DOpE to DOpElib in the documentation, finished overhaul of PDEStatPDE-Examples

In this example, we solve a optimization problems from 
fluid dynamcis. The configuration is similar to 
the fluid optimization problem proposed by Roland Becker 
in 2000 ``Mesh adaption for stationary flow control''.

We describe the underlying equations later...