### 2013-05-11 Examples/OPT/StatPDE/Example5/content.tex
Added some todos

\subsubsection{General problem description}
This example solves the distributed minimization problem
\begin{gather*}
\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
\text{s.t.} (\nabla u,\nabla \phi) = (q+f,\phi)\;\;\forall\,\phi \in H^1_0(\Omega)
\end{gather*}
on the domain $\Omega = [0,1]^2$, and the data is chosen as follows:
\begin{gather*}
 f = \left(20\pi^2  \sin(4 \pi x) - \frac{1}{\alpha}  \sin(\pi x)\right) \sin(2 \pi y)\\
 u^d = \left( 5 \pi^2 \sin(\pi x) + \sin(4 \pi x)\right)  \sin(2\pi y)
\end{gather*}
and $\alpha = 10^{-3}$.
Hence its solution is given by:
\begin{gather*}
 \overline{q} = \frac{1}{\alpha} \sin(\pi x) \sin(2 \pi y)\\
 \overline{u} = \sin(4 \pi x) \sin(2 \pi y)
\end{gather*}

In addition the following functionals are evaluated:
\begin{align*}
  &\text{MidPoint: } u_h(0.5 ; 0.5)&&
  \text{ L1-Value: }\int_\Omega |u_h|\\
  &\text{QError: }\int_\Omega |q_h-\overline{q}|^2
  &&\text{  UError: }\int_\Omega |u_h-\overline{u}|^2
\end{align*}

The important new feature is that we can now use two different meshes for control and state variable.
This is tested first for globally refined meshes, and then for locally refined meshes with different refinements 
for the control and state variable.