### 2013-05-11 Examples/OPT/StatPDE/Example6/content.tex
Example 6 done, changes to localnsontraintacc

\subsubsection{General problem description}
1This example implements the minimum compliance problem for the thickness optimization 
of an MBB-Beam. Using the MMA-Method of K. Svanberg together with an augmented Lagrangian 
approach for the subproblems following M. Stingl.
%\todo{Gleichung, reference}