### 2013-05-11 Examples/OPT/StatPDE/Example6/content.tex
Added doxygen comments to opt_algorithms. Moved mma_algorithm and related files to OPT/StatPDE/Examples6 which is the only example where this is useful.

\subsubsection{General problem description}
1This example implements the minimum compliance problem for the thickness optimization 
of an MBB-Beam. Using the MMA-Method of K. Svanberg together with an augmented Lagrangian 
approach for the subproblems following M. Stingl.
%\todo{Gleichung, reference}

The implementation is done using the following three additional files:

\begin{itemize}
\item \texttt{generalized\underline{ }mma\underline{ }algorithm.h}
  An implementation of the MMA-Algorithm for structural optimization using an augmented
  Lagrangian formulation for the subproblems. The subproblem is implemented using the 
  special purpose file \texttt{augmentedlagrangianproblem.h}.
\item \texttt{augmentedlagrangianproblem.h} The problem container 
  for the augmented Lagrangian problem.
\item \texttt{voidreducedproblem.h} A wrapper file that eliminates $u$ 
  if it is not present anyways. This is used so that we can use the same routines to 
  solve problems that have no PDE constraint. This is used to fit the augmentedlagrangian 
  problem into our framework.
\end{itemize}