### 2013-07-04 Examples/PDE/InstatPDE/Example6/content.tex
First version of Biot-Lame-Navier problem

\subsubsection{General problem description}

In the present example, we solve the quasi-static 
Biot-Lame-Navier problem.

For literture on this field, we refer to 
Rujie Liu, Mary Wheeler and co-workers from UT Austin.