### 2013-07-05 Examples/PDE/InstatPDE/Example6/content.tex
Update description

\subsubsection{General problem description}

In the present example, we solve the quasi-static 
Biot-Lame-Navier problem. The configuration 
is based on the augmented Mandel problem which 
shows the important Mandel-Cryer effect: First increasing pressure
and then decreasing pressure in time while applying some 
traction force on the top boundary.

For literture on this field, we refer to 
Rujie Liu, Mary Wheeler and co-workers from UT Austin.