### 2013-11-10 Examples/PDE/InstatPDE/Example2/content.tex
Content tex in PDE Instat Examples aufgeraeumt

\subsubsection{General problem description}

In the present example, we solve a 
nonstationary fluid-structure interaction problem in 
arbitrary Lagrangian-Eulerian (ALE) coordinates. 
The underlying equations are stated in the following:
\begin{Problem}[Variational fluid-structure interaction framework]
  \label{eq:fsi:ale:harmonic}
  Find $\{\hat v,\hat u,\hat p\} \in \{ \hat v^D + \hat {\cal V}^0\} 
\times \{ \hat u^D + \hat {\cal V}^0 \}\times \hat {\cal L}$, 
  such that $\hat v (0) = \hat v^0$ and $\hat u(0) = \hat u^0$, for almost all
  time steps $t$, and
  \begin{eqnarray*}
    \begin{aligned}
      (\hat J\hat\rho_f  \partial_t \hat v,\hat\psi^v)_{\hat\Omega_f}  
      +(\hat\rho_f \hat J  (\hat F^{-1}(\hat
      v-\partial_t \hat u)\cdot\hat\nabla) \hat v),
      \hat\psi^v)_{\hat\Omega_f} &\\
      + (\hat J\hat\sigma_f\hat
      F^{-T},\hat\nabla\hat\psi^v)_{\hat\Omega_f}
      %
      + (\hat\rho_s \partial_t \hat v,\hat\psi^v)_{\Omega_s}  
      + (\hat J\hat\sigma_s\hat F^{-T},\hat\nabla\hat\psi^v)_{\hat\Omega_s}&\\
      - \langle \hat g, \hat\psi^v \rangle_{\hat\Gamma_N} -
      (\hat\rho_f \hat J\hat f_f, \hat\psi^v)_{\hat\Omega_f}
      - (\hat\rho_s\hat f_s, \hat\psi^v)_{\hat\Omega_s}
      &=0&&\forall\hat\psi^v\in \hat V^0,
      \\
      %%%%%%%%%%%% 
      (\partial_t\hat u-\hat v,\hat\psi^u)_{\hat\Omega_s} 
      + 
      (\hat\sigma_g ,\hat \nabla\hat\psi^u)_{\hat\Omega_f}
      -\langle \hat\sigma_g \hat n_f ,\hat\psi^u\rangle_{\hat \Gamma_i}
      &=0&&\forall\hat\psi^u\in \hat V^0,\\
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
      (\widehat{\text{div}}\,(\hat J\hat F^{-1}
      \hat v_f),\hat\psi^p)_{\hat\Omega_f} 
      + (\hat p_s ,\hat \psi^p)_{\hat\Omega_s}
      &=0&&\forall\hat\psi^p\in \hat L,
    \end{aligned}
  \end{eqnarray*}  
  with the densities $\hat\rho_f$ and $\hat\rho_s$, 
the viscosity $\nu_f$, the Lam\'e parameters 
$\mu_s$, $\lambda_s$ and the deformation gradient $\hat F$, and its
determinant $\hat J$. The stress tensors for the fluid and structure are
implemented by 
\[
\hat\sigma_f = -\hat pI + \hat\rho_f\nu_f (\hat\nabla\hat v \hat F^{-1}
  + \hat F^{-T} \hat\nabla\hat v^T)
\], 
and 
\[
\hat\sigma_s = \hat F (\lambda_s \text{tr}\hat E I + 2\mu_s \hat E)
\]
with $\hat E = \frac{1}{2}(\hat F^T \hat F - I)$.
The mesh deformation
is of harmonic type and given by 
\[
\hat\sigma_g = \alpha(x) \hat\nabla\hat u,
\] 
with a specific parameter $\alpha(x)$ (motivation by Stein/Tezduyar/Benney 2003).
\end{Problem}
The ALE approach belongs to interface-tracking methods in which 
the mesh is moved such that it fits in all time steps with 
the FSI-interface. However, this leads to 
a degeneration of the ALE map. Methods to circument such as 
degeneration as long as possible are re-meshing techniques or 
to use a biharmonic mesh motion technique.

\newpage

{\bf Code validation for fluid problems with ALE}

With the ALE code implemented in Example 
\ref{PDE_Stat_FSI_STVK} it is possible to treat fluid 
problems as well as FSI computations. 
In the case of fluid problems the deformation
gradient and its determinant become:
\begin{equation*}
F:= I , \quad \det F = J = 1.
\end{equation*}

The code is validated by the well-known 
fluid- and FSI benchmark problems. Here, 
the results have been summarized in the table below.

 \begin{table}[h]
   \small
   \centering
     \begin{tabular}{llllcccc}    
       \hline
       Configuration & Date         & Refinement & Time dis. & $k$ & $\Delta p$& $F_D$ & $F_L$   \\ \hline\hline
       BFAC 2D-1     & Mar 10, 2010 & 2 global   & BE        & 1.0    & 0.117412  &       &         \\
       BFAC 2D-1     & Mar 10, 2010 & 2 global   & FS        & 1.0    & 0.117412  &       &         \\
       BFAC 2D-1     & Mar 10, 2010 & 2 global   & CN(k)     & 1.0    & 0.117412  &       &         \\ 
       BFAC 2D-1     & Mar 10, 2010 & 2 global   & CN        & 1.0    & 0.117383  &       &         \\ 
       BFAC 2D-2     & Mar 10, 2010 & 2 global   & CN(k)     & 1.0    & 2.31541   &       &         \\ 
       BFAC 2D-2     & Mar 10, 2010 & 3 global   & CN(k)     & 1.0    & 2.32011   &       &         \\   
       BFAC 2D-2     & Mar 12, 2010 & 3 global   & CN(k)     & 1.0e-2 & 2.50288   &       &         \\   
     \end{tabular}
  \end{table}

\subsubsection{Code validation for FSI with ALE}

The results are summarized below:

 \begin{table}[h]
   \small
   \centering
     \begin{tabular}{llllccccc}    
       \hline
       Config. & Date & Refinement & Time dis. & $k$ &  $u_x(A) [\times 10^{-5}]$ &$u_y(A) [\times 10^{-4}]$ & $F_D$ & $F_L$  \\ \hline
       FSI 1   & Mar 12, 2010 & 3 global & BE   & 1.0 & 8.2003  & 2.2732  & &        \\
       FSI 1   & Apr 08, 2010 & 2 global & BE   & 1.0 & 8.2258  & 2.2813  & &        \\
       FSI 1   & Apr 08, 2010 & 2 global & CN(k)& 0.5 & 8.2268  & 2.2813  & &        \\
       FSI 1   & Apr 08, 2010 & 2 global & CN   & 0.5 & 8.2268  & 2.2813  & &        \\
     \end{tabular}
  \end{table}

\vspace{0.2cm}

\textbf{Program description}

The major difference to the first nonstationary program
is the introduction of the 
\begin{verbatim}
void CellTimeEquationExplicit (...) 
\end{verbatim}
to write all time derivative terms explicitely:
\begin{equation*}
(v^n - v^{n-1}, \phi)_{\Omega}.
\end{equation*}
This behavior is useful since (as shown in the above equations)
other solutions variables have to be considered
around $\partial_t v$ such as $J:=J(u)$. 
The same holds for the corresponding matrix part.

\vspace{0.2cm}