### 2013-12-09 Examples/OPT/StatPDE/Example1/content.tex

\subsubsection{General problem description}
\twick{Dieses Beispiel ist nun extremst ausfuehrlich hinsichtlich der
localpde. Da dies ein Manual ist, wuensche ich mir, dass wir dies (respective
moeglicher Rechtschreibfehler und typos) auch so belassen. Mir hat das damals 
bei FSI Opt sehr geholfen.}
This example solves a distributed minimization problem
\twick{and shows how to estimate the error in the cost functional for stationary optimization
problems. The problem reads:}
\begin{gather*}
\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
\text{s.t.} (\nabla u,\nabla \phi) = (q+f,\phi)\;\;\forall\,\phi \in H^1_0(\Omega)
\end{gather*}
on the domain $\Omega = [0,1]^2$, and the data is chosen as follows:
\begin{gather*}
 f = \left(20\pi^2  \sin(4 \pi x) - \frac{1}{\alpha}  \sin(\pi x)\right) \sin(2 \pi y)\\
 u^d = \left( 5 \pi^2 \sin(\pi x) + \sin(4 \pi x)\right)  \sin(2\pi y)
\end{gather*}
and $\alpha = 10^{-3}$.
Hence its solution is given by:
\begin{gather*}
 \overline{q} = \frac{1}{\alpha} \sin(\pi x) \sin(2 \pi y)\\
 \overline{u} = \sin(4 \pi x) \sin(2 \pi y).
\end{gather*}
Thus the exact optimal value of the cost functional can be calculated as 
\begin{gather*}
 J^* = J(\overline{q},\overline{u}) = \frac{1}{8}\Bigl(25\pi^4 + \frac{1}{\alpha}\Bigr).
\end{gather*}

In addition the following functionals are evaluated:
\begin{gather*}
  \text{MidPoint: } u(0.5 ; 0.5)\\[2mm]
  \text{MeanValue: }\int_\Omega u
\end{gather*}

\twick{Ab hier ist alles neu!!}
\subsubsection{Background information and program description}
In the following, we describe all extensions to the previous problems 
relevant to solving PDE-based optimization with \dope{}. So far, 
we had only to implement the \texttt{ElementEquation} and the corresponding
matrix \texttt{ElementMatrix}. Now, based on the idea of the reduced cost
functional, we have to compute certain additional equations representing the 
adjoint, tangent, and adjoint hessian equations
\texttt{ElementEquation\underline{ }U, ElementEquation\underline{ }T,
ElementEquation\underline{ }UTT}
for the state equation and in the same terms arising from the functional
itself. Let us shed some light into all equations by giving some background
information and overview first.

{
In abstract form, we are given the following optimization problem:
\begin{equation*}
J(q,U) \, \rightarrow \, \min , \quad a(q,U)(\Psi) = 0 \quad \forall\Psi V
\end{equation*}
Lagrangian:
\begin{equation*}
{\cal L}(q,U,Z):=J(q,U) - a(q,U)(Z)
\end{equation*}
Optimality system (KKT system):
\begin{align*}
a_U'(q,U)(\Phi , Z) &= J'_U(q,U)(\Phi) \quad \forall \Phi\in V \\
a_q'(q,U)(\chi , Z) &= J'_q(q,U)(\chi) \quad \forall \chi\in Q \\
a(q,U)(\Psi) &= 0 \qquad \forall\Psi\in V
\end{align*}
or equivalently
\begin{align*}
{\cal L}'_U(q,U,Z)(\Phi) &= 0 \quad\forall \Phi\in V \quad\text{(Adjoint Equation)} \\
{\cal L}'_q(q,U,Z)(\chi) &= 0 \quad\forall \chi\in V \quad\text{(Gradient Equation)} \\
{\cal L}'_Z(q,U,Z)(\Psi) &= 0 \quad\forall \Psi\in V \quad\text{(State Equation)}
\end{align*}
The continuous problem is discretized by a standard Galerkin method using 
finite dimensional subspaces $Q_h \times V_h \subset Q\times V$:
\begin{equation*}
J(q,U) \, \rightarrow \, \min , \quad a(q,U)(\Psi) = 0 \quad \forall\Psi V
\end{equation*}
Discrete saddle-point problems
\begin{align*}
a_U'(q_h,U_h)(\Phi_h , Z_h) &= J'_U(q_h,U_h)(\Phi_h) \quad \forall \Phi_h\in V_h \\
a_q'(q_h,U_h)(\chi_h , Z_h) &= J'_q(q_h,U_h)(\chi_h) \quad \forall \chi_h\in Q_h \\
a(q_h,U_h)(\Psi_h) &= 0 \qquad \forall\Psi_h\in V_h
\end{align*}
\subsubsection{Solution process}
In this section, we briefly discuss the solution process for the 
optimization problem. We further details, we refer to the standard literature.
The unconstrained optimal control problem is reformulated as follows. 
We introduce the solution operator $S:Q\rightarrow V$ of the state equation.
Then:
\begin{equation*}
j(q):= J(q,S(q))\, \rightarrow \, \min , \quad a(q,S(q))(\Psi) = 0 \quad 
\forall \Psi \in V.
\end{equation*}
The local existence and sufficient regularity of $S$ is assumed. The 
necessary optimality conditions of first and second order are
\begin{equation*}
j'(q)(\delta q) = 0 , \quad j''(q)(\delta q,\delta q) \geq 0 \quad 
\forall\delta q\in Q.
\end{equation*} 
The derivatives of the reduced functional can be computed using the 
Lagrangian
\begin{equation*}
{\cal L}(q,U,Z) =J(q,U) - a(q,U)(Z)
\end{equation*}
as already introduced. Let $q\in Q$ be given. We derive
 the corresponding state $U=S(q)\in V$ and the dual variable
$Z\in V$. The following terms hold:
}

\textbf{Dual equation}

\begin{equation*}
{\cal L}_U'(q,U,Z)(\Psi) = 0 \quad\forall\Psi\in V
\end{equation*}
Then,
\begin{equation*}
j'(q)(\delta q) = {\cal L}_q' (q,U,Z)(\delta q) \quad \text{for }\delta q\in Q.
\end{equation*}

\textbf{Tangent equation}

For the right-hand side of Newton's method, we also need further terms. Let
$\delta q\in Q$ be given, and let $\delta U\in V$ and $\delta Z\in V$ fulfill
the equation
\begin{equation*}
{\cal L}_{qZ}'' (q,U,Z)(\delta q,\Phi) + {\cal L}_{UZ}'' (q,U,Z)(\delta
U,\Phi)
\quad\forall \Phi\in V,
\end{equation*}
further

\textbf{Dual Hessian equation}

\begin{equation*}
{\cal L}_{qU}'' (q,U,Z)(\delta q,\Phi) + {\cal L}_{UU}'' 
(q,U,Z)(\delta U,\Phi) + {\cal L}_{ZU}'' (q,U,Z)(\delta Z,\Phi)
\quad\forall \Phi\in V,
\end{equation*}
Then, for $\delta r\in Q$, it holds
\begin{align*}
j''(q)(\delta q, \delta r) 
&= {\cal L}_{qq}'' (q,U,Z)(\delta q,\delta r) \\
&+ {\cal L}_{Uq}'' (q,U,Z)(\delta U,\delta r) \\ 
&+ {\cal L}_{Zq}'' (q,U,Z)(\delta Z,\delta r) \quad \text{for }\delta r\in Q.
\end{align*}
These terms build the the Newton scheme as follows:
\begin{align*}
j''(q^n )(\delta q,\chi ) =&\, -j'(q^n)(\chi) \quad\forall\chi \in Q,\\
q^{n+1} :=&\, q^n + \delta q , \quad n=0,1,2,\ldots .
\end{align*}

Finally, we collect the representations of the forgoing work. First, we recall 
the terms of the Lagrangian:

\textbf{Adjoint:} Find $z\in X$ such that
\begin{equation*}
{\cal L}_u' (q,u,z)(\phi) = 0 \quad \forall\phi\in X. 
\end{equation*}
\textbf{Tangent (linearized state):} Find $\delta u\in X$ such that
\begin{equation*}
{\cal L}_{qz}'' (q,u,z)(\delta q, \phi) +
{\cal L}_{uz}'' (q,u,z)(\delta u, \phi) = 0 \quad \forall\phi\in X. 
\end{equation*}
\textbf{Additional adjoint:} Find $\delta z\in X$ such that
\begin{equation*}
{\cal L}_{qu}'' (q,u,z)(\delta q, \phi) +
{\cal L}_{uu}'' (q,u,z)(\delta u, \phi) +
{\cal L}_{zu}'' (q,u,z)(\delta z, \phi) = 0 \quad \forall\phi\in X. 
\end{equation*}

Especially, the equations for stationary optmization problems are
given as follows:

\textbf{Adjoint:} Find $z\in X$ such that
\begin{equation*}
a_u'(q,u)(\phi,z) = J'(u)(\phi) \quad \forall\phi\in X. 
\end{equation*}
\textbf{Tangent:} Find $\delta u\in X$ such that
\begin{equation*}
a_u'(q,u)(\delta u,\phi) = -a_q'(q,u)(\delta q,\phi)  \quad \forall\phi\in X. 
\end{equation*}
\textbf{Additional adjoint:} Find $\delta z\in X$ such that
\begin{equation*}
a_u'(q,u)(\phi,\delta z) = a_{uu}''(q,u)(\delta u,\phi,z)
-a_{qu}''(q,u)(\delta q,\phi,z) + J''(u)(\delta,\phi) \quad \forall\phi\in X. 
\end{equation*}

Using the Lagrangian, the previous expressions read

\textbf{Lagrangian}
\begin{equation*}
{\cal L}(q,u,z) =J(q,u) - a(q,u)(z)
\end{equation*}

\textbf{State Equation}
\begin{align*}
a(q,u)(\phi) = 0 
\end{align*}

\textbf{Dual Equation}

\begin{align*}
{\cal L}_{u}' (q,u,z)(\phi ) &= 0 \\
\Leftrightarrow a_u' (q,u)(\phi,z) &= J_u' (q,u)(\phi)
\end{align*}

\textbf{Tangent Equation}

\begin{align*}
{\cal L}_{qz}'' (q,u,z)(\delta q, \phi) 
+ {\cal L}_{uz}'' (q,u,z)(\delta u, \phi) &= 0 \\
\Leftrightarrow a_u' (q,u)(\delta u,\phi) + a_q' (q,u)(\delta q,\phi) &= 0
\end{align*}

\textbf{Dual Hessian Equation}

\begin{align*}
{\cal L}_{qu}'' (q,u,z)(\delta q, \phi) 
&+ {\cal L}_{uu}'' (q,u,z)(\delta u, \phi) \\
&+ {\cal L}_{zu}'' (q,u,z)(\delta z, \phi) = 0 \\
\Leftrightarrow a_u' (q,u)(\phi, \delta z) 
&+ a_{uu}' (q,u)(\delta u,\phi, z) \\
&+ a_{qu}' (q,u)(\delta q,\phi, z)  = J_{uu}''(q,u)(\cdot ,\cdot )
\end{align*}

\textbf{Newton left hand side: first derivative}

\begin{align*}
j'(q)(\delta q) &= {\cal L}_{q}' (q,u,z)(\delta q) \quad\text{(Gradient equation)} \\
\Leftrightarrow j'(q)(\delta q) &= -a_q' (q,u)(\delta q,z) + J_{q}'(q,u)(\delta q)
\end{align*}

\textbf{Newton right hand side: second derivative}

\begin{align*}
j''(q)(\delta q, \delta r) &= {\cal L}_{qq}'' (q,u,z)(\delta q, \delta r) + 
{\cal L}_{uq}'' (q,u,z)(\delta u, \delta r) \\
&\quad +  {\cal L}_{zq}'' (q,u,z)(\delta z, \delta r) \\
\Leftrightarrow  j''(q)(\delta q, \delta r) &= 
J_{qq}''(q,u)(\delta q, \delta r) -a_{qq}'' (q,u)(\delta q,\delta r, z) \\
&\quad - a_{uq}'' (q,u)(\delta u,\delta r, z) -a_{q}' (q,u)(\delta r,\delta z)  
\end{align*}

\subsubsection{Implementation in \dope{}}

\begin{align*}
\text{ElementEquation (state)} \quad &\Leftrightarrow \quad a(q,u)(\phi) \\
\text{ElementEquationU (adjoint)} \quad 
&\Leftrightarrow \quad a_u'(q,u)(\phi, z) \\ 
\text{ElementEquationUT (tangent)} \quad 
&\Leftrightarrow \quad a_u'(q,u)(\delta u, \phi) \\
\text{ElementEquationUTT (adjoint hessian)} \quad 
&\Leftrightarrow \quad  a_u'(q,u)(\phi, \delta z) \\
\text{ElementEquationQ (gradient)} \quad 
&\Leftrightarrow \quad a_q'(q,u)(\delta q,z) \\
\text{ElementEquationQT (tangent)} \quad 
&\Leftrightarrow \quad a_q'(q,u)(\delta q, \phi) \\
\text{ElementEquationQTT (hessian)} \quad 
&\Leftrightarrow \quad a_{q}'(q,u)(\delta q,\delta z) \\
\text{ElementEquationUU (adjoint hessian)} \quad 
&\Leftrightarrow \quad a_{uu}''(q,u)(\delta u, \phi, z) \\
\text{ElementEquationQU (adjoint hessian)} \quad 
&\Leftrightarrow \quad a_{qu}''(q,u)(\delta q, \phi, z) \\
\text{ElementEquationUQ (hessian)} \quad 
&\Leftrightarrow \quad a_{uq}''(q,u)(\delta u, \delta r, z)\\
\text{ElementEquationQQ (hessian)} \quad 
&\Leftrightarrow \quad a_{qq}''(q,u)(\delta q, \delta r, z)\\
\text{ElementRightHandSide (state)} \quad 
&\Leftrightarrow \quad f(\phi ) \\
\text{ElementMatrix (all)} \quad 
&\Leftrightarrow \quad a'(q,u)(\delta u) \\
\text{ControlElementEquation (gradient or hessian)} \quad 
&\Leftrightarrow \quad {\cal L}_q' (q,u,z)(\delta q) \\
\text{ControlElementMatrix (all)} \quad 
&\Leftrightarrow \quad a(q,u)(\phi)
\end{align*}

Functionals
\begin{align*}
\text{ElementValue (all)} \quad  & \Leftrightarrow \quad J(q,u) \\
\text{ElementValueU (all)} \quad  & \Leftrightarrow \quad J_u'(q,u)(\phi) \\
\text{ElementValueQ (all)} \quad  & \Leftrightarrow \quad J_q'(q,u)(\phi) \\
\text{ElementValueUU (all)} \quad & \Leftrightarrow \quad J_{uu}''(q,u)(\psi, \phi) \\
\text{ElementValueQU (all)} \quad & \Leftrightarrow \quad J_{qu}''(q,u)(\psi, \phi) \\
\text{ElementValueUQ (all)} \quad & \Leftrightarrow \quad J_{uq}''(q,u)(\psi, \phi) \\
\text{ElementValueQQ (all)} \quad & \Leftrightarrow \quad J_{qq}''(q,u)(\psi, \phi) \\
\end{align*}

\subsubsection{Back to the specific equations in this example}
We have
\[
a(q,u)(\phi) = (\nabla u, \nabla\phi) - (q + f, \phi).
\]
and
\[
a_u'(q,u)(\phi, z) = (\nabla\phi, \nabla z)
\]
and
\[
a_u'(q,u)(\delta u, \phi) = (\nabla\delta u, \nabla\phi)
\]
and
\[
a_u'(q,u)(\phi, \delta z) = (\nabla\phi, \nabla\delta z).
\]
and 
\[
a_q'(q,u)(\delta q, \phi) = -(z,\psi^q)
\]
and 
\[
a_q'(q,u)(\delta q, \phi) = (\delta q,\phi)
\]
and finally
\[
a_{q}'(q,u)(\delta q,\delta z) = - (\delta z,\psi^q)
\]
On the functional level, we have the following terms:
\[
J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2d
\]
and 
\[
J_u'(q,u)(\phi) = (u-u^d, \phi)
\]
and
\[
J_q'(q,u)(\phi) = \alpha (q,\psi^q)
\]
and 
\[
J_{uu}''(q,u)(\psi, \phi) = (\delta u, \phi).
\]
All other terms, specifically mixed terms with $QU$ etc. are zero in this
example.

\subsubsection{\texttt{main.cc}}
Finally, the main file of the optimization examples does not look very much 
different than for pure PDE computations - which is one of the crucial aims 
of our library. 
Here, instead of using a \texttt{pdeproblemcontainer}, we use 
now an \texttt{optproblemcontainer}. Furthermore, 
we define \texttt{ReducedNewtonAlgorithm} to solve the optimization problem.
Next, in the body of the main file, we introduce a second FE function for the 
control variable (using the FE\underline{}Nothing function as we will see 
in Example 2 also allows for 0d controls). Then, we define 
a \texttt{COSTFUNCTIONAL}. Finally, the problem is either solved 
by calling \texttt{Alg.Solve(q)} and/or the user might check if the 
derivatives are implemented correctly by calling 
\texttt{Alg.CheckGrads} or \texttt{Alg.CheckHessian}. The latter 
two functionalities are highly recommended to check your initial 
implementation before wondering about your results.