### 2013-12-10 Examples/OPT/StatPDE/Example4/content.tex

\subsubsection{General problem description}
This example solves the minimization problem
\begin{gather*}
\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
\text{s.t. } (\nabla u,\nabla \phi) = (f,\phi)\;\;\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)\\
\end{gather*}
on the domain $\Omega = [0,1]^2$. In addition we set the Dirichlet data of the state on the boundary as follows
\begin{gather*}
 u_0(0,y) = q_0,\quad u_0(1,y) = q_1,\quad u_0(x,0) = q_2,\quad u_0(x,1) = q_3,\\
 u_1 = q_4^3.
\end{gather*}
The data is chosen as follows:
\begin{align*}
 f &= \left(\begin{matrix}20\pi^2  \sin( \pi x) \sin(\pi y)\\1 \end{matrix}\right)\\
 u^d&= \left(\begin{matrix}\sin( \pi x) \sin(\pi y)*x\\x \end{matrix}\right)
\end{align*}
with $\alpha = 10$.
\subsubsection{Program description}
The control in the Dirichlet boundary values is incorporated via the
class \texttt{LocalDirichletData} which is defined in the
file \textit{localdirichletdata.h}. The class is then given to
the \texttt{OptProblemContainer} as a template argument. This is all that is
needed to use the control in the Dirichlet boundary values. 
\twick{
In the main file and the localpde program, we work still with the 
\texttt{FE\underline{}Nothing} element to assign the five controls.
}