### 2013-12-10 Examples/PDE/InstatPDE/Example5/content.tex

\subsubsection{General problem description}

This example differs only slightly from the previous one. Again, we consider the heat equation, this time with an additional nonlinear term
\begin{equation*}
\partial_t u(t,x,y) - \Delta u(t,x,y) + u(t,x,y)^2 = f(t,x,y),
\end{equation*}
but now in two space dimensions and with known solution 
\begin{equation*}
u(t,x,y) = e^{t-t^2} \sin(x) \sin(y).
\end{equation*}
The computational domain is $\Omega \times I = [0,\pi]^2 \times [0,1]$. From the known solution, we can compute the appropriate data 
\begin{align*}
f(t,x,y) &= (3-2t)e^{t-t^2} \sin(x) \sin(y) + e^{(t-t^2)^2} \sin^2(x) \sin^2(y),\\
u_0(x,y) &= \sin(x) \sin(y).
\end{align*}
Furthermore, we have to prescribe homogeneous Dirichlet boundary conditions.

\subsubsection{Program description}

The new feature of this example is the nonhomogeneous right hand side. In examples \ref{PDE_Stat_Laplace_2D} and \ref{PDE_Stat_Laplace_3D}, we regarded stationary problems with nonhomogeneous right hand sides, but up to now, we never involved the time variable into the nonhomogeneity. To do this, \texttt{DOpElib} yields a \texttt{SetTime()} function which has to be applied in the \textit{localpde.h} file as well as at the place where the \texttt{RightHandSideFunction} class is declared (here the \textit{myfunctions.h} file.