### 2013-12-10 Examples/PDE/StatPDE/Example4/content.tex

\subsubsection{General problem description}
In this problem we solve the simple vector valued Laplace equation in 2d
on the unit square $\Omega=[0,1]^2$, i.e. in strong formulation we look for $u=(u_1,u_2)$ s.t.
\begin{align*}
-\Delta u =& f &&\text{in }\Omega.
\end{align*}
We set zero Dirichlet values on $\partial \Gamma$ and choose $f=(1,1)$. The classical example of a PDE.