### 2013-12-10 Examples/OPT/StatPDE/Example8/content.tex

\subsubsection{General problem description}
This example implements the topology optimization of an MBB-Beam given in \\
{\tt OPT/StatPDE/Example7} using the SIMP method.

The solution is computed using the commercial optimization library SNOPT, similar 
to {\tt OPT/StatPDE/Example6} where IPOPT is used. 
This Example demonstrate how global constraints on
the control variable can be included into the optimization call.

\subsubsection{Program description}
The global constraint is inserted in the main file via 
\texttt{constraints(lcc, 1)} instead of 
\texttt{constraints(lcc, 0)} what we had in the previous two examples.