### 2013-12-13 Examples/PDE/StatPDE/Example4/content.tex

\subsubsection{General problem description}
In this problem we solve the simple vector valued Laplace equation in 2d
on the unit square $\Omega=[0,1]^2$, i.e. in strong formulation we look for $u=(u_1,u_2)$ s.t.
\begin{align*}
-\Delta u =& f &&\text{in }\Omega.
\end{align*}
We set zero Dirichlet values on $\partial \Gamma$ and choose $f=(1,1)$. The
classical example of a PDE.

\begin{remark}[Why this example?]
Originally, \dope{} was designed for coupled and nonlinear problems
with possible PDE-based optimization extensions. Later, we decided 
to add the most simplest PDE (the Laplace/Poisson equation) 
in order to demonstrate how DOpE treats
this well-known example. In addition, the first-time user might 
start here to get a feeling for \dope{} and its capabilities.
\end{remark}