### 2015-11-27 Examples/PDE/StatPDE/Example13/content.tex
Started Work on DG-Example (unfinished)

\subsubsection{General problem description}
Within this example, we demonstrate how to use the
dG (discontinuous Galerkin) method for the solution of 
a transport equation. The example corresponds to 
the dealii step-12.

\subsubsection{Implementational Details}