### 2015-12-03 Examples/PDE/InstatPDE/Example7/content.tex
Added Nonstationary Example (isothermal Euler) with dG discretization in time + Bugfixes

\subsubsection{General problem description}
This example solves the isothermal Euler equations
\begin{align*}
 \partial_t(\rho v) + \partial_x (p(\rho)+ \rho v^2) &= - \frac{\lambda}{2D} v|v| - gh' \rho\\
 \partial_t \rho + \partial_x (\rho v) &=0
\end{align*}
on the set $t \in (0,1)$, $x \in (0,2)$.
The relation 
\[
p(\rho) = \frac{RT\, \rho}{1-\alpha RT\,\rho}
\]
depends on the user provided data for gas-constant $R$ and temperature $T$ as well as a parameter 
$\alpha$ ($\alpha = 0$ for ideal gases).
The other parameters in the system to be provided by the user are 
the friction parameter $\lambda$, the diameter $D$ of the pipe, the gravity $g$ and the slope $h'$ of the 
pipe.

The discretization is done by a dG-method in space.