### 2016-05-13 Examples/OPT/StatPDE/Example10/content.tex

\subsubsection{General problem description}
This example is similar to Example~\ref{OPT_Stat_Param_Nonlin_Fluid}.
The notable difference in the setting is that now, the boundary control areas, 
where the fluid can be sucked out of the domain, are in front of the 
circular inclusion. Hense the drag
\[
k(v,p) = \int_{\Gamma_O} n\cdot \sigma(v,p)\cdot d \, ds,
\]
where $\Gamma_O$ denotes the cylinder boundary, and $d$ is a vector in the
direction of the mean in-flow can become negative, i.e., minimizing the 
value of $k(v,p)$ is no longer viable.

Instead, we consider the objective 
\[
K(q,v,p) = |k(v,p)| + \frac{\alpha}{2}||q - q_0||^2
\] 
is to be minimized.

In contrast to all previous examples, this means that the functional can
no longer be calculated by one integration, but instead the values of the 
integration (for the drag) need to be post-processed.