### 2021-07-27 Examples/PDE/InstatPDE/Example10/content.tex
Added PDE/InstatPDE Examples 10, 11, 12

\subsubsection{General problem description}
We consider phase-field fracture propagation on a slit domain, with time dependent 
Dirichlet data and homogeneous Neumann data.
The main novelties are:
\begin{itemize}
\item Imposing a variational inequality constraint ($\frac{\partial \varphi}{\partial t} \leq 0$)
via a Langrage multiplier similar to  \cite{mang2020phase};
%\item Implementation of an augmented Lagrangian loop (not yet)
\item Stress splitting into tension and compression \`a
la \cite{MieWelHof10a}, see as well \cite{AmorMarigoMaurini2009}, located 
in 
\begin{verbatim}
stress_splitting.h
\end{verbatim}
\item Implementation of a Newton solver that 
allows for an increase of the residual for which, however,
no convergence results are available:
\begin{verbatim}
instat_step_modified_newtonsolver.h
\end{verbatim}
\item The localpde* files:
\begin{verbatim}
localpde_quasi_monolithic.h
\end{verbatim}
which provides the implementation of a quasi-monolithic scheme
(similar to\cite{HeWheWi15}).
\end{itemize}

The problem formulation as well as the handling of the irreversibility constraint is exactly the same as in Example 5.2.8.

The test case is the so-called single edged notched 
tension test. For further details about this problem it is referred
to \cite{MieHofWel2010,BoVeScoHuLa12}, and \cite{WeWiWo2014}.

% \subsubsection{Future improvements}
% Possible future improvements are:
% \begin{itemize}
% \item Implement the phase-field extrapolation in the displacement equation proposed in \cite{HeWheWi15};
% \item Implement the augmented Lagrangian loop proposed by
% Wheeler/Wick/Wollner, CMAME 2014 \cite{WeWiWo2014}. 
% \end{itemize}

% Need to implement extrapolation with two 
% previous time steps
% To be changed: 
% instatpdeproblem.h
% instat_step_newton_solver.h
%
% augmented Lagrangian solver loop
% -> currently simple penalization
% has to be implemented in instat_step_newton_solver.h
%
% Dokumentation in content.tex