### 2021-07-27 Examples/OPT/StatPDE/Example8/content.tex
Added Benchmark fracture examples - and spellcheck in content files

\subsubsection{General problem description}
This example implements the topology optimization of an MBB-Beam given in \\
\texttt{OPT/StatPDE/Example7} using the SIMP method.

The solution is computed using the commercial optimization library SNOPT, similar 
to \texttt{OPT/StatPDE/Example6} where IPOPT is used. 
This Example demonstrate how global constraints on
the control variable can be included into the optimization call.

\subsubsection{Program description}
In this example, we have now some local (point) constraints, local constraints and a global
constraint. In contrast to the previous example, these 
constraints are all taken care of in the 
\texttt{localconstraints.h} file. 
Specifically, the point constraints are handled by the 
\texttt{DOpE::PointConstraints<...>} function, which is also initialized 
in the main file.