### 2022-07-11 Examples/PDE/StatPDE/Example12/content.tex
Now the spacetimehandlers utilize the userprovided component-block mapping also in the renumbering of the dofs

\subsubsection{General problem description}
This example shows the use of the Raviart-Thomas (RT) element in \texttt{DOpElib}. The example 
is taken from \texttt{dealii} step-20 and shows how this step can be implemented in \texttt{DOpElib}. The vector valued Laplace equation is solved in the mixed formulation.

Most things are identical to all the other programs. However there is a subtle difference in 
\texttt{localpde.h}: The block\underline{ }component vector mapping
blocks to components has 3 entries, but since the first finite element
is non-primitive we need to assign both components (0 and 1) to the
same block.

The second difference is that we have to initialize the mapping explicitly. This is due to the fact that the 
default 
\begin{verbatim}
  DOpEWrapper::Mapping<DIM,DOFHANDLER> mapping(1,false);
\end{verbatim}
is not working with the RT-element and leads to elements on which the divergence is NaN.

An additional feature of this example is that for the solution of the PDE in mixed form we are using the 
Schur complement solver provided in \texttt{dealii} step-20. Hence this example shows how simple it is to 
use self-made linear solvers within the \texttt{DOpElib} framework.